---
title: "SA_E1A — Stage 4: Hypothesis Testing"
author: "David G. Kamper"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
```

# Overview

This document implements the seven preregistered hypotheses for SA_E1A, following
the analysis plan exactly as specified in the AsPredicted preregistration. All tests
are at α = .05 unless otherwise noted. No family-wise correction is applied across
the hypothesis set, as justified in the preregistration (the hypotheses address
conceptually distinct questions).

**Important limitation:** The perceived-importance question (`istrue`) was only
presented to participants who selected the conditional or neither interpretation
(n = 14). This limits H3 (n = 6 for the "absolutely essential" subgroup) and
H6 (n = 14 for bimodality testing). These analyses are reported as preregistered
but should be interpreted with extreme caution given the small samples.


# 1. Setup

```{r config}
# ──────────────────────────────────────────────────────────────────────
# PATHS
# ──────────────────────────────────────────────────────────────────────

data_dir <- "/Users/dgkamper/Library/CloudStorage/GoogleDrive-dgkamper@gmail.com/My Drive/DGK Lab/Collaborations/BlankLang Lab/DGK Lab - BlankLangLab - Second Amendment Language Dependencies/Data/Study 1"

analytic_file        <- file.path(data_dir, "SA_E1A_analytic.csv")
classification_file  <- file.path(data_dir, "SA_E1A_rewrite_classifications.csv")
```

```{r libraries, message=FALSE}
# ──────────────────────────────────────────────────────────────────────
# LIBRARIES
#
# lme4:    Generalized and linear mixed-effects models (H1, H2, H7)
# mclust:  Gaussian mixture models (H6)
# boot:    Bootstrap confidence intervals (H5)
# emmeans: Estimated marginal means and contrasts (H2)
# tidyr:   Data reshaping (H2 long format)
# ──────────────────────────────────────────────────────────────────────

library(lme4)
library(mclust)
library(boot)
library(emmeans)
library(tidyr)
```

```{r load-data}
# ──────────────────────────────────────────────────────────────────────
# LOAD DATA
# ──────────────────────────────────────────────────────────────────────

df <- read.csv(analytic_file, stringsAsFactors = FALSE)
cat("Analytic sample:", nrow(df), "participants\n")
cat("Sentences represented:", length(unique(df$sentence_id)), "of 20\n")

# Load rewrite classifications from Stage 3
rewrite_codes <- read.csv(classification_file, stringsAsFactors = FALSE)
df <- merge(df, rewrite_codes[, c("prolific_pid", "rewrite_code")],
            by = "prolific_pid", all.x = TRUE)
cat("Rewrite classifications joined:", sum(!is.na(df$rewrite_code)), "of", nrow(df), "\n")
```

```{r prepare-variables}
# ──────────────────────────────────────────────────────────────────────
# PREPARE VARIABLES
#
# Convert response columns to appropriate types and create derived
# variables needed for hypothesis testing.
# ──────────────────────────────────────────────────────────────────────

# Core DVs — convert from character to numeric
df$interpretation <- as.integer(df$interpretation)  # 1=conditional, 2=categorical, 3=neither
df$confidence     <- as.integer(df$confidence)       # 1-7 scale
df$remove         <- as.integer(df$remove)           # 1-7 scale (1=same, 4=midpoint, 7=different)
df$comma_3v2      <- as.integer(df$comma_3v2)        # 1-4 scale
df$comma_3v1      <- as.integer(df$comma_3v1)        # 1-4 scale
df$comma_2v1      <- as.integer(df$comma_2v1)        # 1-4 scale
df$sentence_id    <- as.factor(df$sentence_id)       # factor for random effects

# Binary interpretation variable for H1
# Exclude "neither" (interpretation = 3) from the primary binary analysis.
# Categorical = 1, Conditional = 0.
df$interp_binary <- ifelse(df$interpretation == 2, 1,
                    ifelse(df$interpretation == 1, 0, NA))

# Perceived importance (istrue): only available for conditional/neither interpreters
# 1 = not crucial, 2 = somewhat important, 3 = absolutely essential
df$istrue <- as.integer(ifelse(df$istrue == "" | is.na(df$istrue), NA, df$istrue))

# Interpretation group labels (for readability in output)
df$interp_label <- ifelse(df$interpretation == 2, "categorical",
                   ifelse(df$interpretation == 1, "conditional", "neither"))

cat("\nVariable summary:\n")
cat("  Interpretation: categorical =", sum(df$interpretation == 2),
    ", conditional =", sum(df$interpretation == 1),
    ", neither =", sum(df$interpretation == 3), "\n")
cat("  Participants with istrue data:", sum(!is.na(df$istrue)), "\n")
cat("  Rewrite codes:", table(df$rewrite_code), "\n")
```


# 2. H1 — Categorical Interpretation Bias

**Preregistered prediction:** Participants will predominantly interpret the sentence
as categorical rather than conditional.

**Test:** Logistic mixed-effects model with sentence-level random intercept.
`interp_binary ~ 1 + (1 | sentence_id)`, where categorical = 1 and conditional = 0.
A significant positive intercept indicates a bias toward categorical interpretation.
"Neither" responses are excluded from this binary analysis but reported descriptively.

```{r h1}
# ──────────────────────────────────────────────────────────────────────
# H1: CATEGORICAL INTERPRETATION BIAS
# ──────────────────────────────────────────────────────────────────────

# Subset to binary responses only (exclude "neither")
df_binary <- df[!is.na(df$interp_binary), ]
cat("H1 sample: n =", nrow(df_binary), "(excluding", sum(df$interpretation == 3), "'neither' responses)\n\n")

# Fit logistic GLMM with bobyqa optimizer as preregistered
h1_model <- glmer(
  interp_binary ~ 1 + (1 | sentence_id),
  data = df_binary,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa")
)

# Extract results
h1_summary <- summary(h1_model)
h1_intercept <- fixef(h1_model)["(Intercept)"]
h1_se <- sqrt(vcov(h1_model)["(Intercept)", "(Intercept)"])
h1_z <- h1_intercept / h1_se
h1_p <- 2 * pnorm(abs(h1_z), lower.tail = FALSE)

# Convert log-odds to probability
h1_prob <- plogis(h1_intercept)

cat("═══ H1 RESULTS ═══\n")
cat("  Log-odds intercept:", round(h1_intercept, 3), "\n")
cat("  SE:", round(h1_se, 3), "\n")
cat("  z:", round(h1_z, 3), "\n")
cat("  p:", format.pval(h1_p, digits = 3), "\n")
cat("  Predicted probability of categorical:", round(h1_prob, 3), "\n")
cat("  Observed proportion categorical:", round(mean(df_binary$interp_binary), 3), "\n")
cat("\n  'Neither' responses (excluded):", sum(df$interpretation == 3),
    "(", round(mean(df$interpretation == 3) * 100, 1), "%)\n")

# Print full model summary for reference
print(h1_summary)
```


# 3. H2 — Comma Sensitivity

**Preregistered prediction:** The 3-comma vs. 1-comma comparison (two commas removed)
will receive higher meaning-change ratings than the one-comma-removed comparisons.

**Test:** Linear mixed model on the three within-subject comma ratings. Each participant
provides three ratings (comma_3v2, comma_3v1, comma_2v1). We reshape to long format
and fit: `rating ~ comparison + (1 | prolific_pid) + (1 | sentence_id)`.
Primary contrast: 3v1 vs. average of (3v2 and 2v1).

```{r h2}
# ──────────────────────────────────────────────────────────────────────
# H2: COMMA SENSITIVITY
# ──────────────────────────────────────────────────────────────────────

# Reshape to long format: three rows per participant
df_comma <- data.frame(
  prolific_pid = rep(df$prolific_pid, 3),
  sentence_id  = rep(df$sentence_id, 3),
  comparison   = factor(rep(c("3v2", "3v1", "2v1"), each = nrow(df)),
                        levels = c("3v2", "3v1", "2v1")),
  rating       = c(df$comma_3v2, df$comma_3v1, df$comma_2v1)
)

cat("H2 long-format data:", nrow(df_comma), "observations",
    "(", nrow(df), "participants × 3 comparisons)\n\n")

# Descriptive means
cat("Descriptive means:\n")
for (comp in c("3v2", "3v1", "2v1")) {
  vals <- df_comma$rating[df_comma$comparison == comp]
  commas_removed <- ifelse(comp == "3v1", "2 commas removed", "1 comma removed")
  cat(sprintf("  %s (%s): mean = %.2f, SD = %.2f\n",
              comp, commas_removed, mean(vals), sd(vals)))
}

# Fit linear mixed model
# Random intercepts for participant and sentence
h2_model <- lmer(
  rating ~ comparison + (1 | prolific_pid) + (1 | sentence_id),
  data = df_comma
)

cat("\nModel summary:\n")
print(summary(h2_model))

# Primary contrast: 3v1 vs. average of (3v2, 2v1)
# Using emmeans for clean pairwise comparisons
h2_emm <- emmeans(h2_model, "comparison")
cat("\nEstimated marginal means:\n")
print(h2_emm)

# Pairwise contrasts
cat("\nPairwise comparisons:\n")
h2_pairs <- pairs(h2_emm)
print(h2_pairs)

# Primary contrast: 3v1 vs. average of one-comma-removed conditions
cat("\nPrimary contrast (3v1 vs. average of 3v2 and 2v1):\n")
h2_contrast <- contrast(h2_emm, list("2-comma vs 1-comma" = c(-0.5, 1, -0.5)))
print(h2_contrast)

cat("\n═══ H2 RESULTS ═══\n")
cat("  Prediction: 3v1 > 3v2 and 3v1 > 2v1\n")
cat("  Observed means: 3v1 =", round(mean(df$comma_3v1), 2),
    ", 3v2 =", round(mean(df$comma_3v2), 2),
    ", 2v1 =", round(mean(df$comma_2v1), 2), "\n")
cat("  Direction consistent with prediction:",
    mean(df$comma_3v1) > mean(df$comma_3v2) &
    mean(df$comma_3v1) > mean(df$comma_2v1), "\n")
```


# 4. H3 — The Contextual Significance Paradox

**Preregistered prediction:** Among participants who rate the prefatory clause as
"absolutely essential" (istrue = 3), mean removal-impact ratings will fall below
the scale midpoint of 4.

**Test:** One-sided t-test. H₀: μ ≥ 4 vs. H₁: μ < 4.

**IMPORTANT LIMITATION:** The perceived-importance question (`istrue`) was only
shown to conditional and neither interpreters in Study 1 (n = 14). Only 6 of these
selected "absolutely essential." This test is therefore severely underpowered.
In the pilot, all participants answered this question.

```{r h3}
# ──────────────────────────────────────────────────────────────────────
# H3: THE PARADOX
# ──────────────────────────────────────────────────────────────────────

# Subset to participants who answered "absolutely essential" (istrue = 3)
df_essential <- df[!is.na(df$istrue) & df$istrue == 3, ]

cat("H3 sample: n =", nrow(df_essential), "participants rated 'absolutely essential'\n")
cat("  (out of", sum(!is.na(df$istrue)), "who answered the perceived-importance question)\n\n")

if (nrow(df_essential) >= 3) {
  # Descriptive
  cat("Removal-impact ratings among 'absolutely essential' raters:\n")
  cat("  Mean:", round(mean(df_essential$remove), 2), "\n")
  cat("  SD:", round(sd(df_essential$remove), 2), "\n")
  cat("  Range: [", min(df_essential$remove), ",", max(df_essential$remove), "]\n\n")
  
  # One-sided t-test: mean < 4 (midpoint)
  h3_test <- t.test(df_essential$remove, mu = 4, alternative = "less")
  
  cat("═══ H3 RESULTS ═══\n")
  cat("  One-sided t-test: H₀: μ ≥ 4 vs H₁: μ < 4\n")
  cat("  t =", round(h3_test$statistic, 3), "\n")
  cat("  df =", round(h3_test$parameter, 1), "\n")
  cat("  p =", format.pval(h3_test$p.value, digits = 3), "\n")
  cat("  95% CI upper bound:", round(h3_test$conf.int[2], 2), "\n")
  cat("  Mean:", round(h3_test$estimate, 2), "\n")
  cat("\n  *** WARNING: n =", nrow(df_essential), "is too small for reliable inference. ***\n")
  cat("  This result should be interpreted with extreme caution.\n")
} else {
  cat("  *** INSUFFICIENT DATA: Cannot run H3 with n =", nrow(df_essential), "***\n")
}
```


# 5. H4 — Confidence Equivalence

**Preregistered prediction:** Both interpretation groups will report confidence above
the scale midpoint (4), and the groups will not differ meaningfully in confidence.

**Test:** (a) Within each group, one-sided t-test: mean confidence > 4.
(b) TOST equivalence test on the group difference with bounds Δ = ±0.5 scale points.

```{r h4}
# ──────────────────────────────────────────────────────────────────────
# H4: CONFIDENCE EQUIVALENCE
# ──────────────────────────────────────────────────────────────────────

df_cat  <- df[df$interpretation == 2, ]
df_cond <- df[df$interpretation == 1, ]

cat("H4 samples: categorical n =", nrow(df_cat), ", conditional n =", nrow(df_cond), "\n\n")

# ── Part (a): one-sided t-tests, mean > 4 ──
cat("H4a: Confidence above midpoint (4)\n\n")

# Categorical interpreters
h4a_cat <- t.test(df_cat$confidence, mu = 4, alternative = "greater")
cat("  Categorical: mean =", round(h4a_cat$estimate, 2),
    ", t =", round(h4a_cat$statistic, 3),
    ", p =", format.pval(h4a_cat$p.value, digits = 3), "\n")

# Conditional interpreters
h4a_cond <- t.test(df_cond$confidence, mu = 4, alternative = "greater")
cat("  Conditional: mean =", round(h4a_cond$estimate, 2),
    ", t =", round(h4a_cond$statistic, 3),
    ", p =", format.pval(h4a_cond$p.value, digits = 3), "\n")

# ── Part (b): TOST equivalence test ──
# The TOST procedure consists of two one-sided t-tests:
#   Test 1: H₀: δ ≤ -Δ  vs  H₁: δ > -Δ  (the difference is not too negative)
#   Test 2: H₀: δ ≥ +Δ  vs  H₁: δ < +Δ  (the difference is not too positive)
# Equivalence is concluded if BOTH tests are significant.

cat("\nH4b: TOST equivalence test (bounds Δ = ±0.5)\n\n")

delta <- 0.5  # equivalence bound in scale points
group_diff <- mean(df_cat$confidence) - mean(df_cond$confidence)

# Welch two-sample t-test infrastructure
# We need the pooled SE of the difference
tost_test <- t.test(df_cat$confidence, df_cond$confidence)
se_diff <- abs(group_diff / tost_test$statistic)  # recover SE from t and difference
df_tost <- tost_test$parameter

# TOST Test 1: difference > -delta (lower bound)
t1 <- (group_diff - (-delta)) / se_diff
p1 <- pt(t1, df = df_tost, lower.tail = FALSE)  # one-sided, testing > -delta

# TOST Test 2: difference < +delta (upper bound)
t2 <- (group_diff - delta) / se_diff
p2 <- pt(t2, df = df_tost, lower.tail = TRUE)   # one-sided, testing < +delta

# The TOST p-value is the larger of the two
tost_p <- max(p1, p2)

cat("  Group difference (cat - cond):", round(group_diff, 2), "\n")
cat("  SE of difference:", round(se_diff, 3), "\n")
cat("  TOST Test 1 (diff > -0.5): t =", round(t1, 3), ", p =", format.pval(p1, digits = 3), "\n")
cat("  TOST Test 2 (diff < +0.5): t =", round(t2, 3), ", p =", format.pval(p2, digits = 3), "\n")
cat("  TOST p-value:", format.pval(tost_p, digits = 3), "\n")
cat("  Equivalence concluded:", tost_p < 0.05, "\n")

cat("\n═══ H4 RESULTS ═══\n")
cat("  H4a: Both groups above midpoint —",
    "Categorical:", h4a_cat$p.value < 0.05,
    ", Conditional:", h4a_cond$p.value < 0.05, "\n")
cat("  H4b: Equivalence (±0.5) concluded:", tost_p < 0.05, "\n")
cat("  90% CI on difference: [", round(group_diff - qt(0.95, df_tost) * se_diff, 2),
    ",", round(group_diff + qt(0.95, df_tost) * se_diff, 2), "]\n")
cat("  NOTE: Conditional subgroup (n =", nrow(df_cond), ") is small;\n")
cat("  this test is underpowered as acknowledged in the preregistration.\n")
```


# 6. H5 — Rewrite Concordance

**Preregistered predictions:**
- H5a (descriptive): Among categorical interpreters, some will produce conditional rewrites.
- H5b (descriptive): Among conditional interpreters, concordance with the rewrite will be imperfect.
- H5c (confirmatory): Concordance rate is higher among categorical than conditional interpreters.

**Rewrites were classified computationally** (blind to forced-choice responses)
using Claude Sonnet via the Anthropic API (Stage 3). The preregistration specifies
"trained raters blind to forced-choice responses"; the computational classifier
serves this function, with human spot-checking as a validation step.

```{r h5}
# ──────────────────────────────────────────────────────────────────────
# H5: REWRITE CONCORDANCE
# ──────────────────────────────────────────────────────────────────────

# Concordance = the forced-choice interpretation matches the rewrite code.
# For categorical FC: concordant if rewrite_code == "categorical"
# For conditional FC: concordant if rewrite_code == "conditional"
# Ambiguous rewrites are neither concordant nor discordant.

df$concordant <- ifelse(
  df$interp_label == "categorical" & df$rewrite_code == "categorical", TRUE,
  ifelse(
    df$interp_label == "conditional" & df$rewrite_code == "conditional", TRUE,
    FALSE
  )
)

# ── Cross-tabulation ──
cat("Forced-Choice × Rewrite Classification:\n\n")
cross_tab <- table(
  "Forced-Choice" = df$interp_label,
  "Rewrite Code" = df$rewrite_code
)
print(cross_tab)

# ── H5a: Discordance among categorical interpreters ──
cat("\n\n═══ H5a: Discordance among categorical interpreters ═══\n")

df_fc_cat <- df[df$interpretation == 2, ]
n_cat_total <- nrow(df_fc_cat)
n_cat_conditional_rewrite <- sum(df_fc_cat$rewrite_code == "conditional")
prop_discordant <- n_cat_conditional_rewrite / n_cat_total

cat("  Categorical FC interpreters:", n_cat_total, "\n")
cat("  Rewrites coded as conditional:", n_cat_conditional_rewrite, "\n")
cat("  Proportion discordant:", round(prop_discordant, 3), "\n")

# Bootstrap CI for the discordant proportion
if (n_cat_conditional_rewrite > 0) {
  boot_fn <- function(data, indices) {
    d <- data[indices]
    mean(d == "conditional")
  }
  set.seed(42)
  boot_h5a <- boot(df_fc_cat$rewrite_code, boot_fn, R = 10000)
  ci_h5a <- boot.ci(boot_h5a, type = "perc")
  cat("  95% bootstrap CI: [", round(ci_h5a$percent[4], 3), ",",
      round(ci_h5a$percent[5], 3), "]\n")
} else {
  # With 0 successes, use exact binomial CI instead
  ci_h5a_exact <- binom.test(0, n_cat_total)$conf.int
  cat("  95% exact binomial CI: [", round(ci_h5a_exact[1], 3), ",",
      round(ci_h5a_exact[2], 3), "]\n")
  cat("  (Exact CI used because proportion = 0)\n")
}
cat("  Pilot comparison: 28.4% discordant\n")


# ── H5b: Concordance among conditional interpreters ──
cat("\n═══ H5b: Concordance among conditional interpreters ═══\n")

df_fc_cond <- df[df$interpretation == 1, ]
n_cond_total <- nrow(df_fc_cond)
n_cond_concordant <- sum(df_fc_cond$rewrite_code == "conditional")
prop_concordant_cond <- n_cond_concordant / n_cond_total

cat("  Conditional FC interpreters:", n_cond_total, "\n")
cat("  Rewrites coded as conditional:", n_cond_concordant, "\n")
cat("  Proportion concordant:", round(prop_concordant_cond, 3), "\n")

if (n_cond_concordant > 0) {
  boot_fn_cond <- function(data, indices) {
    d <- data[indices]
    mean(d == "conditional")
  }
  set.seed(42)
  boot_h5b <- boot(df_fc_cond$rewrite_code, boot_fn_cond, R = 10000)
  ci_h5b <- boot.ci(boot_h5b, type = "perc")
  cat("  95% bootstrap CI: [", round(ci_h5b$percent[4], 3), ",",
      round(ci_h5b$percent[5], 3), "]\n")
} else {
  ci_h5b_exact <- binom.test(0, n_cond_total)$conf.int
  cat("  95% exact binomial CI: [", round(ci_h5b_exact[1], 3), ",",
      round(ci_h5b_exact[2], 3), "]\n")
  cat("  (Exact CI used because proportion = 0)\n")
}
cat("  Pilot comparison: 29.8% concordant\n")


# ── H5c: Concordance comparison ──
cat("\n═══ H5c: Concordance comparison ═══\n")

# Concordance rates (including ambiguous in denominator)
conc_cat  <- sum(df_fc_cat$rewrite_code == "categorical") / n_cat_total
conc_cond <- n_cond_concordant / n_cond_total

cat("  Concordance rate (categorical FC):", round(conc_cat, 3), "\n")
cat("  Concordance rate (conditional FC):", round(conc_cond, 3), "\n")
cat("  Difference:", round(conc_cat - conc_cond, 3), "\n\n")

# Two-sample proportion test (one-sided)
# P(concordance | categorical) > P(concordance | conditional)
h5c_test <- prop.test(
  x = c(sum(df_fc_cat$rewrite_code == "categorical"), n_cond_concordant),
  n = c(n_cat_total, n_cond_total),
  alternative = "greater"
)

cat("  Two-sample proportion test (one-sided):\n")
cat("  χ² =", round(h5c_test$statistic, 3), "\n")
cat("  p =", format.pval(h5c_test$p.value, digits = 3), "\n")
cat("  Pilot comparison: 69.4% vs 29.8% (difference ≈ 40 pp)\n")
```


# 7. H6 — Bimodality of Perceived Importance

**Preregistered prediction:** Responses on the perceived-importance question will be
bimodal, clustering at "not crucial" and "absolutely essential" with few in the middle.

**Test:** Fit Gaussian mixture models with 1 and 2 components using `mclust`.
Lower BIC for the 2-component model constitutes evidence of bimodality. Also fit
a 3-component model as a supplementary check.

**CRITICAL LIMITATION:** Only 14 participants answered this question (conditional
and neither interpreters only). This sample is far too small for reliable Gaussian
mixture modeling. Results are reported for completeness but should not be interpreted
as informative.

```{r h6}
# ──────────────────────────────────────────────────────────────────────
# H6: BIMODALITY OF PERCEIVED IMPORTANCE
# ──────────────────────────────────────────────────────────────────────

df_istrue <- df[!is.na(df$istrue), ]

cat("H6 sample: n =", nrow(df_istrue), "\n\n")

cat("Perceived importance distribution:\n")
istrue_tab <- table(df_istrue$istrue)
istrue_labels <- c("1" = "Not crucial", "2" = "Somewhat important", "3" = "Absolutely essential")
for (val in names(istrue_tab)) {
  cat(sprintf("  %s (%s): %d (%.1f%%)\n",
              val, istrue_labels[val], istrue_tab[val],
              istrue_tab[val] / nrow(df_istrue) * 100))
}

if (nrow(df_istrue) >= 10) {
  # Fit 1-, 2-, and 3-component Gaussian mixture models
  # Adding small noise to discrete responses for mclust compatibility
  set.seed(42)
  istrue_jittered <- df_istrue$istrue + rnorm(nrow(df_istrue), sd = 0.1)
  
  mc1 <- Mclust(istrue_jittered, G = 1)
  mc2 <- Mclust(istrue_jittered, G = 2)
  mc3 <- Mclust(istrue_jittered, G = 3)
  
  cat("\nGaussian mixture model BIC values:\n")
  cat("  1-component:", round(mc1$bic, 2), "\n")
  cat("  2-component:", round(mc2$bic, 2), "\n")
  cat("  3-component:", round(mc3$bic, 2), "\n")
  
  # BIC comparison (higher = better in mclust's convention)
  best <- which.max(c(mc1$bic, mc2$bic, mc3$bic))
  cat("  Best fitting model:", best, "component(s)\n")
  
  cat("\n═══ H6 RESULTS ═══\n")
  cat("  2-component BIC better than 1-component:", mc2$bic > mc1$bic, "\n")
  cat("  *** WARNING: n =", nrow(df_istrue), "is far too small for reliable ***\n")
  cat("  *** mixture model estimation. This result is not informative. ***\n")
} else {
  cat("\n═══ H6 RESULTS ═══\n")
  cat("  *** INSUFFICIENT DATA: n =", nrow(df_istrue), "***\n")
  cat("  Cannot fit Gaussian mixture models with this sample size.\n")
}
```


# 8. H7 — Removal Impact by Interpretation Group

**Preregistered prediction:** Among categorical interpreters, mean removal-impact
will fall significantly below the midpoint of 4. Among conditional interpreters,
the mean will not differ significantly from 4.

**Test:** Categorical: one-sided t-test, H₁: μ < 4. Conditional: two-sided t-test.
A sentence-level random intercept is included if sample size permits; if singular,
fall back to simple t-test.

```{r h7}
# ──────────────────────────────────────────────────────────────────────
# H7: REMOVAL IMPACT BY INTERPRETATION GROUP
# ──────────────────────────────────────────────────────────────────────

# ── Categorical interpreters ──
cat("═══ H7: Removal impact ═══\n\n")
cat("Categorical interpreters (n =", nrow(df_cat), "):\n")
cat("  Mean removal:", round(mean(df_cat$remove), 2), "\n")
cat("  SD:", round(sd(df_cat$remove), 2), "\n\n")

# Try mixed model first, fall back to t-test if singular
h7_cat_mixed <- tryCatch({
  m <- lmer(remove ~ 1 + (1 | sentence_id), data = df_cat)
  # Check for singular fit
  if (isSingular(m)) {
    cat("  Mixed model: singular fit, falling back to t-test.\n")
    NULL
  } else {
    cat("  Mixed model: converged with sentence random intercept.\n")
    m
  }
}, error = function(e) {
  cat("  Mixed model failed:", e$message, "\n  Falling back to t-test.\n")
  NULL
})

if (!is.null(h7_cat_mixed)) {
  # Use the mixed model estimate
  h7_cat_est <- fixef(h7_cat_mixed)["(Intercept)"]
  h7_cat_se <- sqrt(vcov(h7_cat_mixed)["(Intercept)", "(Intercept)"])
  h7_cat_t <- (h7_cat_est - 4) / h7_cat_se
  # Approximate df using Satterthwaite or just use large-sample z
  h7_cat_p <- pnorm(h7_cat_t)  # one-sided, testing < 4
  cat("  Intercept:", round(h7_cat_est, 2), ", SE:", round(h7_cat_se, 2), "\n")
  cat("  Test against midpoint (4): z =", round(h7_cat_t, 3), ", p =", format.pval(h7_cat_p, digits = 3), "\n")
} else {
  # Simple one-sided t-test
  h7_cat_t <- t.test(df_cat$remove, mu = 4, alternative = "less")
  cat("  t =", round(h7_cat_t$statistic, 3), "\n")
  cat("  df =", round(h7_cat_t$parameter, 1), "\n")
  cat("  p =", format.pval(h7_cat_t$p.value, digits = 3), "\n")
}

# ── Conditional interpreters ──
cat("\n\nConditional interpreters (n =", nrow(df_cond), "):\n")
cat("  Mean removal:", round(mean(df_cond$remove), 2), "\n")
cat("  SD:", round(sd(df_cond$remove), 2), "\n\n")

# With n = 9, a mixed model with sentence random intercepts likely won't converge.
# Try it, but expect to fall back to t-test.
h7_cond_mixed <- tryCatch({
  m <- lmer(remove ~ 1 + (1 | sentence_id), data = df_cond)
  if (isSingular(m)) {
    cat("  Mixed model: singular fit, falling back to t-test.\n")
    NULL
  } else {
    cat("  Mixed model: converged with sentence random intercept.\n")
    m
  }
}, error = function(e) {
  cat("  Mixed model failed:", e$message, "\n  Falling back to t-test.\n")
  NULL
})

if (!is.null(h7_cond_mixed)) {
  h7_cond_est <- fixef(h7_cond_mixed)["(Intercept)"]
  h7_cond_se <- sqrt(vcov(h7_cond_mixed)["(Intercept)", "(Intercept)"])
  h7_cond_t <- (h7_cond_est - 4) / h7_cond_se
  h7_cond_p <- 2 * pnorm(-abs(h7_cond_t))  # two-sided
  cat("  Intercept:", round(h7_cond_est, 2), ", SE:", round(h7_cond_se, 2), "\n")
  cat("  Test against midpoint (4): z =", round(h7_cond_t, 3), ", p =", format.pval(h7_cond_p, digits = 3), "\n")
} else {
  h7_cond_t <- t.test(df_cond$remove, mu = 4, alternative = "two.sided")
  cat("  t =", round(h7_cond_t$statistic, 3), "\n")
  cat("  df =", round(h7_cond_t$parameter, 1), "\n")
  cat("  p =", format.pval(h7_cond_t$p.value, digits = 3), "\n")
}

cat("\n═══ H7 SUMMARY ═══\n")
cat("  Categorical mean removal:", round(mean(df_cat$remove), 2), "(predicted < 4)\n")
cat("  Conditional mean removal:", round(mean(df_cond$remove), 2), "(predicted ≈ 4)\n")
```


# 9. Summary of All Hypotheses

```{r summary-table}
# ──────────────────────────────────────────────────────────────────────
# COMPREHENSIVE SUMMARY
# ──────────────────────────────────────────────────────────────────────

cat("═══════════════════════════════════════════════════════════════════\n")
cat("  SA_E1A HYPOTHESIS TESTING — COMPLETE SUMMARY\n")
cat("  Analytic N = ", nrow(df), "\n")
cat("═══════════════════════════════════════════════════════════════════\n\n")

cat("  H1  Categorical bias:     ", round(mean(df_binary$interp_binary) * 100, 1),
    "% categorical (pilot: 84.3%)\n")
cat("  H2  Comma sensitivity:    3v1 mean =", round(mean(df$comma_3v1), 2),
    "vs 3v2 =", round(mean(df$comma_3v2), 2),
    ", 2v1 =", round(mean(df$comma_2v1), 2), "\n")
cat("  H3  Paradox:              n = 6 'essential' raters, mean remove =",
    ifelse(nrow(df_essential) > 0, round(mean(df_essential$remove), 2), "N/A"),
    " (UNDERPOWERED)\n")
cat("  H4  Confidence equiv:     Cat =", round(mean(df_cat$confidence), 2),
    ", Cond =", round(mean(df_cond$confidence), 2), "\n")
cat("  H5a Discordance (cat FC): ", n_cat_conditional_rewrite, "/",
    n_cat_total, "conditional rewrites\n")
cat("  H5b Concordance (cond FC):", n_cond_concordant, "/",
    n_cond_total, "conditional rewrites\n")
cat("  H5c Concordance diff:     ", round(conc_cat, 3), "vs",
    round(conc_cond, 3), "\n")
cat("  H6  Bimodality:           n =", sum(!is.na(df$istrue)), "(UNDERPOWERED)\n")
cat("  H7  Removal by group:     Cat =", round(mean(df_cat$remove), 2),
    ", Cond =", round(mean(df_cond$remove), 2), "\n")

cat("\n  Key limitation: istrue only collected for conditional/neither\n")
cat("  interpreters (n = 14), severely limiting H3 and H6.\n")
cat("═══════════════════════════════════════════════════════════════════\n")
```


# 10. Additional Measures: "And" Insertion and Distinguish

These measures were collected from all participants but are not tied to specific
preregistered hypotheses. They are reported descriptively and broken down by
interpretation group.

## 10.1 "And" Insertion Impact

Participants saw the original sentence alongside a version with "and" inserted
before the operative clause (e.g., "...being necessary for X, **and** the right
of people to Y, must not be Z"). They rated the difference on the same 1-7
scale as the removal question (1 = same implications, 7 = completely different).

```{r and-insertion}
# ──────────────────────────────────────────────────────────────────────
# "AND" INSERTION IMPACT
# ──────────────────────────────────────────────────────────────────────

df$and_impact <- as.integer(df$and)

cat("═══ 'And' Insertion Impact ═══\n\n")

# Overall descriptives
cat("Overall:\n")
cat("  Mean:", round(mean(df$and_impact), 2), "\n")
cat("  SD:", round(sd(df$and_impact), 2), "\n")
cat("  Median:", median(df$and_impact), "\n")
cat("  Distribution:\n")
and_tab <- table(df$and_impact)
for (val in names(and_tab)) {
  cat(sprintf("    %s: %d (%.1f%%)\n", val, and_tab[val],
              and_tab[val] / length(df$and_impact) * 100))
}

# Test against midpoint (4): is insertion perceived as changing meaning?
and_overall_t <- t.test(df$and_impact, mu = 4)
cat("\n  Test against midpoint (4):\n")
cat("    t =", round(and_overall_t$statistic, 3),
    ", df =", round(and_overall_t$parameter, 1),
    ", p =", format.pval(and_overall_t$p.value, digits = 3), "\n")
cat("    Mean is", ifelse(mean(df$and_impact) < 4, "below", "above"), "midpoint\n")

# By interpretation group
cat("\nBy interpretation group:\n")
for (grp in c("categorical", "conditional", "neither")) {
  sub <- df[df$interp_label == grp, ]
  if (nrow(sub) > 0) {
    cat(sprintf("  %s (n=%d): M=%.2f, SD=%.2f\n",
                grp, nrow(sub), mean(sub$and_impact), sd(sub$and_impact)))
  }
}

# Compare categorical vs conditional (if enough data)
if (sum(df$interpretation == 1) >= 3) {
  and_group_t <- t.test(df$and_impact[df$interpretation == 2],
                         df$and_impact[df$interpretation == 1])
  cat("\n  Cat vs. Cond t-test: t =", round(and_group_t$statistic, 3),
      ", p =", format.pval(and_group_t$p.value, digits = 3), "\n")
}
```

## 10.2 Distinguish Question

Participants chose which phrase best replaces "being necessary for" in the
sentence. Options: (1) "for the purpose of," (2) "because," or (3) "neither."
This captures whether participants read the prefatory clause as expressing
a purpose relationship or a causal/factual one.

```{r distinguish}
# ──────────────────────────────────────────────────────────────────────
# DISTINGUISH QUESTION
# ──────────────────────────────────────────────────────────────────────

df$distinguish <- as.integer(df$distinguish)

dist_labels <- c("1" = "For the purpose of",
                 "2" = "Because",
                 "3" = "Neither")

cat("═══ Distinguish Question ═══\n\n")

# Overall distribution
cat("Overall distribution:\n")
dist_tab <- table(df$distinguish)
for (val in names(dist_tab)) {
  cat(sprintf("  %s (%s): %d (%.1f%%)\n",
              val, dist_labels[val], dist_tab[val],
              dist_tab[val] / nrow(df) * 100))
}

# By interpretation group
cat("\nBy interpretation group:\n")
for (grp in c("categorical", "conditional", "neither")) {
  sub <- df[df$interp_label == grp, ]
  if (nrow(sub) > 0) {
    cat(sprintf("\n  %s (n=%d):\n", grp, nrow(sub)))
    grp_tab <- table(sub$distinguish)
    for (val in names(grp_tab)) {
      cat(sprintf("    %s (%s): %d (%.1f%%)\n",
                  val, dist_labels[val], grp_tab[val],
                  grp_tab[val] / nrow(sub) * 100))
    }
  }
}

# Chi-square test: does distinguish choice differ by interpretation group?
# Use only categorical vs. conditional (exclude "neither" group due to small n)
dist_cross <- table(
  df$interp_label[df$interpretation %in% c(1, 2)],
  df$distinguish[df$interpretation %in% c(1, 2)]
)
cat("\n\nCross-tabulation (categorical vs conditional only):\n")
print(dist_cross)

# Fisher's exact test (some cells may be small)
dist_fisher <- fisher.test(dist_cross)
cat("\n  Fisher's exact test: p =", format.pval(dist_fisher$p.value, digits = 3), "\n")
```

## 10.3 Comma Insertion (AddCommas)

Participants saw their sentence with all commas removed and were asked to
add commas wherever they thought they should go. The original sentences have
three commas at canonical positions:

- **Position 1:** After the subject, before "being necessary"
- **Position 2:** After the purpose clause, before "the right of"
- **Position 3:** After the action phrase, before "must not be" / "shall not be"

We detect each position with regex and classify each response by its
comma placement pattern.

```{r addcommas-positions}
# ──────────────────────────────────────────────────────────────────────
# COMMA INSERTION ANALYSIS
#
# Participants copied the sentence (stripped of commas) and added
# commas where they thought they belonged. We check whether commas
# appear at each of the three canonical positions.
# ──────────────────────────────────────────────────────────────────────

cat("═══ Comma Insertion (AddCommas) ═══\n\n")

# Detect comma at each canonical position
df$comma_p1 <- grepl(",\\s*being\\s+necessary", df$addcommas, ignore.case = TRUE)
df$comma_p2 <- grepl(",\\s*the right of", df$addcommas, ignore.case = TRUE)
df$comma_p3 <- grepl(",\\s*(must|shall)\\s+not", df$addcommas, ignore.case = TRUE)

# Total commas placed
df$n_commas_placed <- nchar(gsub("[^,]", "", df$addcommas))

# Pattern string: e.g., "110" means positions 1 and 2 have commas, 3 does not
df$comma_pattern <- paste0(
  ifelse(df$comma_p1, "1", "0"),
  ifelse(df$comma_p2, "1", "0"),
  ifelse(df$comma_p3, "1", "0")
)

# ── How many commas did participants place? ──
cat("Total commas placed:\n")
nc_tab <- table(df$n_commas_placed)
for (val in names(nc_tab)) {
  cat(sprintf("  %s commas: %d (%.1f%%)\n", val, nc_tab[val],
              nc_tab[val] / nrow(df) * 100))
}
cat(sprintf("\n  Mean: %.2f, SD: %.2f\n", mean(df$n_commas_placed),
            sd(df$n_commas_placed)))

# ── Position-by-position rates ──
cat("\nComma placement at each canonical position:\n")
cat(sprintf("  Position 1 (before 'being necessary'):  %d/100 (%d%%)\n",
            sum(df$comma_p1), sum(df$comma_p1)))
cat(sprintf("  Position 2 (before 'the right of'):     %d/100 (%d%%)\n",
            sum(df$comma_p2), sum(df$comma_p2)))
cat(sprintf("  Position 3 (before 'must/shall not'):   %d/100 (%d%%)\n",
            sum(df$comma_p3), sum(df$comma_p3)))

# ── Pattern frequencies ──
cat("\nPlacement patterns (1 = comma present, 0 = absent):\n")
cat("  Pattern  P1  P2  P3  Count  Description\n")
pat_tab <- sort(table(df$comma_pattern), decreasing = TRUE)
pat_desc <- c(
  "111" = "All three (matches original)",
  "110" = "Subject + purpose (dropped action comma)",
  "010" = "Purpose boundary only",
  "011" = "Purpose + action (dropped subject comma)",
  "101" = "Subject + action (dropped purpose comma)",
  "100" = "Subject only",
  "001" = "Action only",
  "000" = "No commas at canonical positions"
)
for (pat in names(pat_tab)) {
  desc <- ifelse(pat %in% names(pat_desc), pat_desc[pat], "Other")
  cat(sprintf("  %s     %s   %s   %s   %3d    %s\n",
              pat, substr(pat,1,1), substr(pat,2,2), substr(pat,3,3),
              pat_tab[pat], desc))
}
```

```{r addcommas-by-group}
# ──────────────────────────────────────────────────────────────────────
# COMMA PLACEMENT BY INTERPRETATION GROUP
#
# Does comma placement pattern differ between categorical and
# conditional interpreters? If conditional interpreters are more
# sensitive to the clause boundaries, they might place more commas
# or place them at different positions.
# ──────────────────────────────────────────────────────────────────────

cat("\n═══ Comma Placement by Interpretation Group ═══\n\n")

# Mean commas by group
cat("Mean commas placed:\n")
for (grp in c("categorical", "conditional", "neither")) {
  sub <- df[df$interp_label == grp, ]
  if (nrow(sub) > 0) {
    cat(sprintf("  %s (n=%d): M=%.2f, SD=%.2f\n",
                grp, nrow(sub), mean(sub$n_commas_placed),
                sd(sub$n_commas_placed)))
  }
}

# Position rates by group
cat("\nPosition rates by group (% placing comma at each position):\n")
cat("  Group           P1 (subject)  P2 (purpose)  P3 (action)\n")
for (grp in c("categorical", "conditional", "neither")) {
  sub <- df[df$interp_label == grp, ]
  if (nrow(sub) > 0) {
    cat(sprintf("  %-14s  %5.1f%%        %5.1f%%        %5.1f%%\n",
                grp,
                mean(sub$comma_p1) * 100,
                mean(sub$comma_p2) * 100,
                mean(sub$comma_p3) * 100))
  }
}

# Pattern cross-tab (categorical vs conditional only)
cat("\nPattern × Interpretation (categorical vs conditional):\n")
pat_cross <- table(
  df$interp_label[df$interpretation %in% c(1, 2)],
  df$comma_pattern[df$interpretation %in% c(1, 2)]
)
print(pat_cross)

# Did they recover the original 3-comma structure?
cat("\nProportion recovering original 3-comma structure (pattern 111):\n")
for (grp in c("categorical", "conditional", "neither")) {
  sub <- df[df$interp_label == grp, ]
  if (nrow(sub) > 0) {
    pct <- mean(sub$comma_pattern == "111") * 100
    cat(sprintf("  %s: %.1f%% (%d/%d)\n", grp, pct,
                sum(sub$comma_pattern == "111"), nrow(sub)))
  }
}
```

```{r addcommas-position3}
# ──────────────────────────────────────────────────────────────────────
# POSITION 3 IS THE MOST VARIABLE
#
# Position 3 (the comma before "must not be") is the most contested,
# with only 50% of participants placing it. This is the joint between
# the operative clause's content (the right to do X) and its verb
# (must not be restricted). Dropping this comma fuses the two into
# a single clause. Does this relate to interpretation?
# ──────────────────────────────────────────────────────────────────────

cat("\n═══ Position 3 and Interpretation ═══\n\n")

# Is Position 3 placement associated with interpretation?
# Among binary interpreters only
df_p3 <- df[df$interpretation %in% c(1, 2), ]

cat("Position 3 comma × Interpretation:\n")
p3_cross <- table(
  "Interpretation" = df_p3$interp_label,
  "Comma at P3" = df_p3$comma_p3
)
print(p3_cross)

# Fisher's exact test
p3_fisher <- fisher.test(p3_cross)
cat("\n  Fisher's exact test: p =", format.pval(p3_fisher$p.value, digits = 3), "\n")

# Position 3 and removal impact: do people who drop the comma
# also perceive removal of the prefatory clause as less impactful?
cat("\nRemoval impact by Position 3 comma placement:\n")
cat(sprintf("  With P3 comma (n=%d):    M=%.2f, SD=%.2f\n",
            sum(df$comma_p3), mean(df$remove[df$comma_p3]),
            sd(df$remove[df$comma_p3])))
cat(sprintf("  Without P3 comma (n=%d): M=%.2f, SD=%.2f\n",
            sum(!df$comma_p3), mean(df$remove[!df$comma_p3]),
            sd(df$remove[!df$comma_p3])))
p3_rem_t <- t.test(df$remove[df$comma_p3], df$remove[!df$comma_p3])
cat(sprintf("  t = %.3f, p = %s\n",
            p3_rem_t$statistic, format.pval(p3_rem_t$p.value, digits = 3)))
```


# 11. Demographics

Descriptive statistics for all demographic variables collected. These are
reported for sample characterization; none were preregistered as predictors.

```{r demographics}
# ──────────────────────────────────────────────────────────────────────
# DEMOGRAPHICS
# ──────────────────────────────────────────────────────────────────────

cat("═══ Demographics ═══\n\n")

# ── Gender ──
# 1 = male, 2 = female, 3 = other/non-binary
gender_labels <- c("1" = "Male", "2" = "Female", "3" = "Other/Non-binary")
cat("Gender:\n")
gen_tab <- table(df$gender)
for (val in names(gen_tab)) {
  cat(sprintf("  %s: %d (%.1f%%)\n",
              gender_labels[val], gen_tab[val],
              gen_tab[val] / nrow(df) * 100))
}

# ── Age ──
df$age_num <- as.numeric(df$age)
cat("\nAge:\n")
cat("  Mean:", round(mean(df$age_num, na.rm = TRUE), 1), "\n")
cat("  SD:", round(sd(df$age_num, na.rm = TRUE), 1), "\n")
cat("  Range:", min(df$age_num, na.rm = TRUE), "-", max(df$age_num, na.rm = TRUE), "\n")
cat("  Median:", median(df$age_num, na.rm = TRUE), "\n")

# ── Race ──
# 1=White, 2=Black, 3=Asian, 4=Hispanic/Latino, 5=Native American,
# 6=Other, multi-select separated by semicolons
cat("\nRace/Ethnicity (including multi-select):\n")
race_tab <- table(df$race)
race_labels <- c("1" = "White", "2" = "Black", "3" = "Asian",
                 "4" = "Hispanic/Latino", "5" = "Native American",
                 "6" = "Other")
for (val in names(sort(race_tab, decreasing = TRUE))) {
  # Try to decode multi-select entries
  decoded <- paste(sapply(strsplit(val, ";")[[1]], function(x) {
    ifelse(x %in% names(race_labels), race_labels[x], x)
  }), collapse = " + ")
  cat(sprintf("  %s: %d (%.1f%%)\n",
              decoded, race_tab[val], race_tab[val] / nrow(df) * 100))
}
```


# 12. Gun Control Attitudes (Exploratory)

Participants answered: "In general, do you think gun laws should be made more
strict, less strict, or kept as they are now?" on a 5-point scale:
1 = Much more strict, 2 = Somewhat more strict, 3 = Kept as they are now,
4 = Somewhat less strict, 5 = Much less strict.

This analysis is entirely exploratory. The key question is whether gun control
attitudes predict interpretation, but given the 90.5% categorical rate, there
is very little variance in the outcome to explain.

## 12.1 Gun Control Descriptives

```{r gun-descriptives}
# ──────────────────────────────────────────────────────────────────────
# GUN CONTROL ATTITUDES — DESCRIPTIVES
# ──────────────────────────────────────────────────────────────────────

df$gun_control <- as.integer(df$gun_control)

gun_labels <- c("1" = "Much more strict",
                "2" = "Somewhat more strict",
                "3" = "Kept as they are",
                "4" = "Somewhat less strict",
                "5" = "Much less strict")

cat("═══ Gun Control Attitudes ═══\n\n")

cat("Distribution:\n")
gun_tab <- table(df$gun_control)
for (val in names(gun_tab)) {
  cat(sprintf("  %s (%s): %d (%.1f%%)\n",
              val, gun_labels[val], gun_tab[val],
              gun_tab[val] / nrow(df) * 100))
}
cat(sprintf("\n  Mean: %.2f, SD: %.2f, Median: %d\n",
            mean(df$gun_control), sd(df$gun_control), median(df$gun_control)))
cat("  (Lower values = more pro-restriction)\n")
```

## 12.2 Gun Control and Interpretation

```{r gun-interpretation}
# ──────────────────────────────────────────────────────────────────────
# GUN CONTROL × INTERPRETATION
#
# Does gun control attitude predict whether a participant chooses
# categorical vs. conditional? Given the 90.5% categorical rate,
# we expect little variance to work with. We report both a logistic
# regression (treating gun_control as continuous) and a descriptive
# breakdown.
# ──────────────────────────────────────────────────────────────────────

cat("═══ Gun Control × Interpretation ═══\n\n")

# Descriptive: mean gun control by interpretation group
cat("Mean gun control attitude by interpretation group:\n")
for (grp in c("categorical", "conditional", "neither")) {
  sub <- df[df$interp_label == grp, ]
  if (nrow(sub) > 0) {
    cat(sprintf("  %s (n=%d): M=%.2f, SD=%.2f\n",
                grp, nrow(sub), mean(sub$gun_control), sd(sub$gun_control)))
  }
}

# Full distribution cross-tab
cat("\nInterpretation × Gun control cross-tabulation:\n")
cross_gun <- table(df$interp_label, df$gun_control)
print(cross_gun)

# Logistic regression: gun_control predicting binary interpretation
# (excluding "neither" responses)
df_binary_gun <- df[!is.na(df$interp_binary), ]

# Simple logistic regression (no random effects, since gun_control is
# a participant-level predictor and we have one observation per participant)
gun_logit <- glm(interp_binary ~ gun_control, data = df_binary_gun,
                  family = binomial)
cat("\n\nLogistic regression: interpretation ~ gun_control\n")
print(summary(gun_logit))

gun_coef <- coef(gun_logit)["gun_control"]
gun_se <- summary(gun_logit)$coefficients["gun_control", "Std. Error"]
gun_p <- summary(gun_logit)$coefficients["gun_control", "Pr(>|z|)"]
gun_or <- exp(gun_coef)

cat(sprintf("\n  Gun control coefficient: %.3f (SE = %.3f, p = %s)\n",
            gun_coef, gun_se, format.pval(gun_p, digits = 3)))
cat(sprintf("  Odds ratio: %.3f\n", gun_or))
cat("  (OR > 1 means more pro-gun = more likely categorical;\n")
cat("   OR < 1 means more pro-gun = less likely categorical)\n")
```

## 12.3 Gun Control and Confidence

```{r gun-confidence}
# ──────────────────────────────────────────────────────────────────────
# GUN CONTROL × CONFIDENCE
# ──────────────────────────────────────────────────────────────────────

cat("═══ Gun Control × Confidence ═══\n\n")

gun_conf_cor <- cor.test(df$gun_control, df$confidence)
cat("  Pearson r =", round(gun_conf_cor$estimate, 3), "\n")
cat("  p =", format.pval(gun_conf_cor$p.value, digits = 3), "\n")
cat("  95% CI: [", round(gun_conf_cor$conf.int[1], 3), ",",
    round(gun_conf_cor$conf.int[2], 3), "]\n")
```

## 12.4 Gun Control and Removal Impact

```{r gun-removal}
# ──────────────────────────────────────────────────────────────────────
# GUN CONTROL × REMOVAL IMPACT
# ──────────────────────────────────────────────────────────────────────

cat("═══ Gun Control × Removal Impact ═══\n\n")

gun_rem_cor <- cor.test(df$gun_control, df$remove)
cat("  Pearson r =", round(gun_rem_cor$estimate, 3), "\n")
cat("  p =", format.pval(gun_rem_cor$p.value, digits = 3), "\n")
cat("  95% CI: [", round(gun_rem_cor$conf.int[1], 3), ",",
    round(gun_rem_cor$conf.int[2], 3), "]\n")
```

## 12.5 Gun Control and Rewrite Concordance

```{r gun-rewrite}
# ──────────────────────────────────────────────────────────────────────
# GUN CONTROL × REWRITE CODE
#
# Does gun control attitude relate to what kind of rewrite people
# produce? With zero conditional rewrites, this is really asking
# whether gun attitudes predict categorical vs. ambiguous rewrites.
# ──────────────────────────────────────────────────────────────────────

cat("═══ Gun Control × Rewrite Code ═══\n\n")

cat("Mean gun control by rewrite classification:\n")
for (code in c("categorical", "ambiguous", "conditional")) {
  sub <- df[df$rewrite_code == code, ]
  if (nrow(sub) > 0) {
    cat(sprintf("  %s (n=%d): M=%.2f, SD=%.2f\n",
                code, nrow(sub), mean(sub$gun_control), sd(sub$gun_control)))
  }
}
```

## 12.6 Gun Control and "And" Insertion Impact

```{r gun-and}
# ──────────────────────────────────────────────────────────────────────
# GUN CONTROL × "AND" INSERTION IMPACT
# ──────────────────────────────────────────────────────────────────────

cat("═══ Gun Control × 'And' Insertion Impact ═══\n\n")

gun_and_cor <- cor.test(df$gun_control, df$and_impact)
cat("  Pearson r =", round(gun_and_cor$estimate, 3), "\n")
cat("  p =", format.pval(gun_and_cor$p.value, digits = 3), "\n")
cat("  95% CI: [", round(gun_and_cor$conf.int[1], 3), ",",
    round(gun_and_cor$conf.int[2], 3), "]\n")
```

## 12.7 Gun Control and Distinguish

```{r gun-distinguish}
# ──────────────────────────────────────────────────────────────────────
# GUN CONTROL × DISTINGUISH
#
# Does gun control attitude relate to whether participants interpret
# "being necessary" as purposive ("for the purpose of") or causal
# ("because")?
# ──────────────────────────────────────────────────────────────────────

cat("═══ Gun Control × Distinguish ═══\n\n")

cat("Mean gun control by distinguish choice:\n")
for (val in c(1, 2, 3)) {
  sub <- df[df$distinguish == val, ]
  if (nrow(sub) > 0) {
    label <- c("1" = "For the purpose of", "2" = "Because", "3" = "Neither")[as.character(val)]
    cat(sprintf("  %s (n=%d): M=%.2f, SD=%.2f\n",
                label, nrow(sub), mean(sub$gun_control), sd(sub$gun_control)))
  }
}

# One-way ANOVA
gun_dist_aov <- summary(aov(gun_control ~ factor(distinguish), data = df))
cat(sprintf("\n  One-way ANOVA: F = %.3f, p = %s\n",
            gun_dist_aov[[1]]["factor(distinguish)", "F value"],
            format.pval(gun_dist_aov[[1]]["factor(distinguish)", "Pr(>F)"], digits = 3)))
```

## 12.8 Gun Control and Comma Placement

```{r gun-commas-placed}
# ──────────────────────────────────────────────────────────────────────
# GUN CONTROL × COMMA PLACEMENT
#
# Do gun attitudes predict how many commas participants place, or
# whether they place a comma at the contested Position 3?
# ──────────────────────────────────────────────────────────────────────

cat("═══ Gun Control × Comma Placement ═══\n\n")

# Number of commas placed
gun_ncomma_cor <- cor.test(df$gun_control, df$n_commas_placed)
cat("Gun control × number of commas placed:\n")
cat("  Pearson r =", round(gun_ncomma_cor$estimate, 3),
    ", p =", format.pval(gun_ncomma_cor$p.value, digits = 3), "\n")

# Position 3 specifically
cat("\nGun control × Position 3 placement:\n")
cat(sprintf("  With P3 comma (n=%d): M gun = %.2f\n",
            sum(df$comma_p3), mean(df$gun_control[df$comma_p3])))
cat(sprintf("  Without P3 comma (n=%d): M gun = %.2f\n",
            sum(!df$comma_p3), mean(df$gun_control[!df$comma_p3])))
gun_p3_t <- t.test(df$gun_control[df$comma_p3], df$gun_control[!df$comma_p3])
cat(sprintf("  t = %.3f, p = %s\n",
            gun_p3_t$statistic, format.pval(gun_p3_t$p.value, digits = 3)))
```

## 12.9 Gun Control and Comma Sensitivity Ratings

```{r gun-comma-ratings}
# ──────────────────────────────────────────────────────────────────────
# GUN CONTROL × COMMA SENSITIVITY RATINGS (the H2 measures)
# ──────────────────────────────────────────────────────────────────────

cat("═══ Gun Control × Comma Sensitivity Ratings ═══\n\n")

for (comp in c("comma_3v2", "comma_3v1", "comma_2v1")) {
  r <- cor.test(df$gun_control, df[[comp]])
  cat(sprintf("  %s: r = %.3f, p = %s\n",
              comp, r$estimate, format.pval(r$p.value, digits = 3)))
}
```


# 13. Age and Gender as Predictors (Exploratory)

```{r age-gender}
# ──────────────────────────────────────────────────────────────────────
# AGE AND GENDER × ALL DVs
# ──────────────────────────────────────────────────────────────────────

cat("═══ Age × Interpretation ═══\n\n")

# Age by interpretation group
for (grp in c("categorical", "conditional", "neither")) {
  sub <- df[df$interp_label == grp, ]
  if (nrow(sub) > 0) {
    cat(sprintf("  %s (n=%d): M=%.1f, SD=%.1f, range=[%d, %d]\n",
                grp, nrow(sub),
                mean(sub$age_num, na.rm = TRUE),
                sd(sub$age_num, na.rm = TRUE),
                min(sub$age_num, na.rm = TRUE),
                max(sub$age_num, na.rm = TRUE)))
  }
}

# Logistic regression: age predicting interpretation (binary)
age_logit <- glm(interp_binary ~ age_num, data = df_binary_gun, family = binomial)
age_p <- summary(age_logit)$coefficients["age_num", "Pr(>|z|)"]
cat(sprintf("\n  Logistic regression (interpretation ~ age): p = %s\n",
            format.pval(age_p, digits = 3)))

# Age × other DVs
cat("\n═══ Age × Other DVs ═══\n\n")
for (dv_name in c("confidence", "remove", "and_impact", "n_commas_placed")) {
  r <- cor.test(df$age_num, df[[dv_name]], use = "complete.obs")
  cat(sprintf("  Age × %s: r = %.3f, p = %s\n",
              dv_name, r$estimate, format.pval(r$p.value, digits = 3)))
}

# Age × comma sensitivity
cat("\n")
for (comp in c("comma_3v2", "comma_3v1", "comma_2v1")) {
  r <- cor.test(df$age_num, df[[comp]], use = "complete.obs")
  cat(sprintf("  Age × %s: r = %.3f, p = %s\n",
              comp, r$estimate, format.pval(r$p.value, digits = 3)))
}

cat("\n\n═══ Gender × Interpretation ═══\n\n")

# Gender by interpretation group
df$gender_label <- ifelse(df$gender == "1", "Male",
                   ifelse(df$gender == "2", "Female", "Other"))

gender_interp <- table(df$gender_label, df$interp_label)
cat("Gender × Interpretation:\n")
print(gender_interp)

# Proportions within each gender
cat("\nProportion categorical by gender:\n")
for (g in c("Male", "Female")) {
  sub <- df[df$gender_label == g, ]
  pct_cat <- mean(sub$interpretation == 2) * 100
  cat(sprintf("  %s (n=%d): %.1f%% categorical\n", g, nrow(sub), pct_cat))
}

# Fisher's exact test (male vs female only, exclude other due to n=1)
gender_binary <- df[df$gender_label %in% c("Male", "Female") &
                     df$interpretation %in% c(1, 2), ]
gender_fisher <- fisher.test(table(gender_binary$gender_label,
                                    gender_binary$interp_label))
cat(sprintf("\n  Fisher's exact (Male vs Female, cat vs cond): p = %s\n",
            format.pval(gender_fisher$p.value, digits = 3)))

# Gender × other DVs
cat("\n═══ Gender × Other DVs (Male vs Female) ═══\n\n")
df_mf <- df[df$gender_label %in% c("Male", "Female"), ]
for (dv_name in c("confidence", "remove", "and_impact", "n_commas_placed")) {
  tt <- t.test(df_mf[[dv_name]][df_mf$gender_label == "Male"],
               df_mf[[dv_name]][df_mf$gender_label == "Female"])
  cat(sprintf("  %s: Male M=%.2f, Female M=%.2f, t=%.3f, p=%s\n",
              dv_name,
              mean(df_mf[[dv_name]][df_mf$gender_label == "Male"]),
              mean(df_mf[[dv_name]][df_mf$gender_label == "Female"]),
              tt$statistic, format.pval(tt$p.value, digits = 3)))
}
```


# 14. Comprehensive Summary (Updated)

```{r summary-updated}
# ──────────────────────────────────────────────────────────────────────
# FULL SUMMARY INCLUDING EXPLORATORY ANALYSES
# ──────────────────────────────────────────────────────────────────────

cat("═══════════════════════════════════════════════════════════════════\n")
cat("  SA_E1A COMPLETE ANALYSIS SUMMARY\n")
cat("  Analytic N = ", nrow(df), "\n")
cat("═══════════════════════════════════════════════════════════════════\n\n")

cat("PREREGISTERED HYPOTHESES\n")
cat("  H1  Categorical bias:      ", round(mean(df_binary$interp_binary) * 100, 1),
    "% categorical\n")
cat("  H2  Comma sensitivity:     3v1 =", round(mean(df$comma_3v1), 2),
    ", 3v2 =", round(mean(df$comma_3v2), 2),
    ", 2v1 =", round(mean(df$comma_2v1), 2), "\n")
cat("  H3  Paradox:               n = 6, UNDERPOWERED\n")
cat("  H4  Confidence:            Cat =", round(mean(df_cat$confidence), 2),
    ", Cond =", round(mean(df_cond$confidence), 2),
    "; equivalence NOT established\n")
cat("  H5  Concordance:           0 conditional rewrites (from anyone)\n")
cat("  H6  Bimodality:            n = 14, UNDERPOWERED\n")
cat("  H7  Removal:               Cat =", round(mean(df_cat$remove), 2),
    " (< midpoint), Cond =", round(mean(df_cond$remove), 2), "\n")

cat("\nADDITIONAL MEASURES\n")
cat("  And insertion:             M =", round(mean(df$and_impact), 2),
    " (below midpoint)\n")
cat("  Distinguish:               ",
    round(table(df$distinguish)["2"] / nrow(df) * 100, 1),
    "% chose 'because'\n")
cat("  Comma insertion:           ",
    sum(df$comma_pattern == "111"), "% recovered all 3 commas;",
    "P1=", paste0(sum(df$comma_p1), "%"),
    "P2=", paste0(sum(df$comma_p2), "%"),
    "P3=", paste0(sum(df$comma_p3), "%"), "\n")

cat("\nEXPLORATORY (NOT PREREGISTERED)\n")
cat("  Gun control distribution:  ",
    sum(df$gun_control <= 2), "pro-restriction,",
    sum(df$gun_control == 3), "neutral,",
    sum(df$gun_control >= 4), "anti-restriction\n")
cat("  Gun control x interp:      p =", format.pval(gun_p, digits = 3), "\n")
cat("  Age x interp:              p =", format.pval(age_p, digits = 3), "\n")
cat("  Gender x interp:           p =",
    format.pval(gender_fisher$p.value, digits = 3), "\n")

cat("\n  Sample: ", sum(df$gender == "1"), "male,",
    sum(df$gender == "2"), "female;",
    "mean age =", round(mean(df$age_num, na.rm = TRUE), 1), "\n")
cat("═══════════════════════════════════════════════════════════════════\n")
```
