---
title: "SA_E1A — Stage 2: Exclusion Criteria"
author: "David G. Kamper"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
```

# Overview

This document applies the preregistered exclusion criteria to the merged dataset
produced in Stage 1. Exclusions are applied **sequentially** (each step operates
on the participants remaining after the previous step), and we report the number
excluded at each step, as specified in the preregistration.

The preregistered exclusion criteria are:

1. AI detection task (image captcha)
2. Self-paced reading task (also for AI detection)
3. Rewrite quality (Levenshtein edit distance)
4. Self-reported non-seriousness
5. Self-reported AI use
6. Non-consent or device failure (handled by experiment logic)
7. Total completion time (under 3 minutes)

We also exclude one test run (PID = "TEST") that was used for platform verification
before data collection began.

The output is a filtered CSV: **`SA_E1A_analytic.csv`**.


# 1. Setup

```{r config}
# ──────────────────────────────────────────────────────────────────────
# PATHS
# Input: the merged CSV from Stage 1
# Output: the analytic sample after exclusions, saved in the same dir
# ──────────────────────────────────────────────────────────────────────

data_dir <- "/Users/dgkamper/Library/CloudStorage/GoogleDrive-dgkamper@gmail.com/My Drive/DGK Lab/Collaborations/BlankLang Lab/DGK Lab - BlankLangLab - Second Amendment Language Dependencies/Data/Study 1"

input_file  <- file.path(data_dir, "SA_E1A_merged.csv")
output_file <- file.path(data_dir, "SA_E1A_analytic.csv")
```

```{r libraries}
library(stringdist)  # for stringdist() — Levenshtein edit distance
```

```{r load-data}
# ──────────────────────────────────────────────────────────────────────
# LOAD MERGED DATA
# This is the output of Stage 1: one row per participant, with
# standardized column names and sentence metadata already joined.
# ──────────────────────────────────────────────────────────────────────

df <- read.csv(input_file, stringsAsFactors = FALSE)
cat("Loaded:", nrow(df), "participants.\n")
```


# 2. Exclusion Tracking

We create a tracking data frame to record whether each participant is excluded
and at which step, so we can report a transparent exclusion cascade.

```{r exclusion-setup}
# ──────────────────────────────────────────────────────────────────────
# EXCLUSION TRACKER
# Each participant starts as "included". As we apply each criterion,
# we mark participants who fail. Because exclusions are sequential,
# a participant is only counted under the FIRST criterion they fail.
# (If someone fails both captcha and seriousness, they're counted
# under captcha since that comes first.)
# ──────────────────────────────────────────────────────────────────────

df$excluded       <- FALSE     # will be set to TRUE if any criterion fails
df$exclusion_step <- NA_character_  # which step excluded them (for reporting)
n_start <- nrow(df)

# Helper function: apply one exclusion criterion
# - condition: logical vector (TRUE = should be excluded)
# - label: human-readable label for reporting
# Returns the number newly excluded at this step
apply_exclusion <- function(condition, label) {
  
  # Only apply to participants not already excluded
  newly_excluded <- condition & !df$excluded
  n_new <- sum(newly_excluded)
  
  # Mark them
  df$excluded[newly_excluded]       <<- TRUE
  df$exclusion_step[newly_excluded] <<- label
  
  # Report
  n_remaining <- sum(!df$excluded)
  cat(sprintf("  %-45s  excluded: %d   remaining: %d\n", 
              label, n_new, n_remaining))
  
  return(n_new)
}

cat("Starting N:", n_start, "\n\n")
```


# 3. Apply Exclusion Criteria

```{r step-0}
# ──────────────────────────────────────────────────────────────────────
# STEP 0: Test / platform-verification runs
#
# Before the study launched, the experimenter completed test runs to
# verify the Pavlovia + Prolific integration. These are identified by
# placeholder Prolific PIDs (e.g., "TEST"). They are not real
# participants and must be removed before applying substantive
# exclusion criteria.
#
# This step is not part of the preregistered exclusion sequence (it
# precedes it), since test runs are not participant data.
# ──────────────────────────────────────────────────────────────────────

cat("Step 0: Test runs\n")
apply_exclusion(
  condition = df$prolific_pid %in% c("TEST", "test", ""),
  label = "0. Test/verification run"
)
```

```{r step-1}
# ──────────────────────────────────────────────────────────────────────
# STEP 1: AI detection — image captcha
#
# The experiment displays an image instructing participants to press
# the letter "Q". Hidden white-on-white HTML text says "press Z" to
# trap automated browsers (which read HTML, not images). Participants
# who press Z (or any key other than Q) fail. The variable
# captcha_pass is TRUE for Q, FALSE for Z.
#
# Participants who fail are ejected by the experiment itself (they see
# a "Thank you" screen and cannot continue). So in practice, anyone
# in our dataset who completed the experiment should have
# captcha_pass = TRUE. We check anyway for safety.
#
# Note: The TEST run (excluded above) has captcha_pass = NA because
# it predates the captcha being added to the experiment.
# ──────────────────────────────────────────────────────────────────────

cat("Step 1: AI detection — captcha\n")
apply_exclusion(
  condition = is.na(df$captcha_pass) | df$captcha_pass != "true",
  label = "1. Failed captcha"
)
```

```{r step-2}
# ──────────────────────────────────────────────────────────────────────
# STEP 2: AI detection — self-paced reading
#
# Participants complete two moving-window self-paced reading (SPR)
# trials as a secondary AI detection measure. Automated systems would
# produce implausibly fast or perfectly uniform reading times.
#
# The preregistration states that the SPR task is "also for AI
# detection" and that initial data collection was used "to calibrate
# exclusion criteria for self-paced reading time thresholds."
#
# Criterion: We exclude participants whose minimum word-level reading
# time on EITHER SPR trial falls below 100 milliseconds. Human
# readers cannot extract word meaning in under ~100ms (cf. Rayner,
# 1998; typical lower bounds in SPR studies are 100–150ms). Times
# below this threshold indicate automated spacebar pressing.
#
# Note: SPR data is missing (NA) for the TEST run (already excluded
# in Step 0). For remaining participants, we require non-missing SPR
# data.
# ──────────────────────────────────────────────────────────────────────

cat("Step 2: AI detection — self-paced reading\n")

# Convert to numeric (they are stored as character from CSV)
df$spr1_min_rt <- as.numeric(df$spr1_min_rt)
df$spr2_min_rt <- as.numeric(df$spr2_min_rt)

# Criterion: min RT below 100ms on either trial, or missing SPR data
# for a participant who wasn't already excluded
spr_fail <- (!df$excluded) & (
  is.na(df$spr1_min_rt) | is.na(df$spr2_min_rt) |
  df$spr1_min_rt < 100 | df$spr2_min_rt < 100
)

apply_exclusion(
  condition = spr_fail,
  label = "2. Implausible SPR times (<100ms)"
)

# Descriptive: show SPR distributions for remaining participants
remaining <- df[!df$excluded, ]
cat(sprintf("\n  SPR1 min RT — median: %.0f ms, range: [%.0f, %.0f]\n",
            median(remaining$spr1_min_rt, na.rm = TRUE),
            min(remaining$spr1_min_rt, na.rm = TRUE),
            max(remaining$spr1_min_rt, na.rm = TRUE)))
cat(sprintf("  SPR2 min RT — median: %.0f ms, range: [%.0f, %.0f]\n",
            median(remaining$spr2_min_rt, na.rm = TRUE),
            min(remaining$spr2_min_rt, na.rm = TRUE),
            max(remaining$spr2_min_rt, na.rm = TRUE)))
```

```{r step-3}
# ──────────────────────────────────────────────────────────────────────
# STEP 3: Rewrite quality — Levenshtein edit distance
#
# For the Rewrite question, participants were asked to rewrite the
# sentence "in your own words, making it as simple as possible."
# Near-verbatim copies indicate the participant did not engage with
# the task meaningfully (they may have copy-pasted despite the paste
# prevention, or typed it out verbatim).
#
# Preregistered criterion: Compute the Levenshtein edit distance
# between the original sentence and the participant's rewrite AFTER
# removing all non-alphabetic characters and converting to lowercase.
# Exclude participants with edit distance < 5.
#
# The stringdist package computes Levenshtein distance efficiently.
# ──────────────────────────────────────────────────────────────────────

cat("Step 3: Rewrite quality — edit distance\n")

# Clean both strings: remove non-alpha characters, lowercase
clean_text <- function(x) {
  tolower(gsub("[^a-zA-Z]", "", x))
}

df$rewrite_clean  <- clean_text(df$rewrite)
df$sentence_clean <- clean_text(df$full_sentence)

# Compute Levenshtein distance
df$edit_distance <- stringdist(df$rewrite_clean, df$sentence_clean, method = "lv")

# Apply criterion
apply_exclusion(
  condition = !is.na(df$edit_distance) & df$edit_distance < 5,
  label = "3. Near-verbatim rewrite (edit dist < 5)"
)

# Descriptive: show the excluded rewrites for transparency
excluded_rewrites <- df[df$exclusion_step == "3. Near-verbatim rewrite (edit dist < 5)" & 
                          !is.na(df$exclusion_step), ]
if (nrow(excluded_rewrites) > 0) {
  cat("\n  Excluded rewrites:\n")
  for (i in seq_len(nrow(excluded_rewrites))) {
    cat(sprintf("    PID=%s, SID=%d, edit_dist=%d\n", 
                excluded_rewrites$prolific_pid[i],
                excluded_rewrites$sentence_id[i],
                excluded_rewrites$edit_distance[i]))
    cat(sprintf("      Original: %s\n", substr(excluded_rewrites$full_sentence[i], 1, 80)))
    cat(sprintf("      Rewrite:  %s\n\n", substr(excluded_rewrites$rewrite[i], 1, 80)))
  }
}

# Show the distribution of edit distances for remaining participants
remaining <- df[!df$excluded, ]
cat(sprintf("  Edit distance — median: %.0f, range: [%.0f, %.0f]\n",
            median(remaining$edit_distance, na.rm = TRUE),
            min(remaining$edit_distance, na.rm = TRUE),
            max(remaining$edit_distance, na.rm = TRUE)))
```

```{r step-4}
# ──────────────────────────────────────────────────────────────────────
# STEP 4: Self-reported non-seriousness
#
# In the debrief, participants were asked: "Did you complete this
# experiment seriously throughout (without responding randomly)?"
# Response options:
#   1 = "No, I didn't"
#   4 = "Yes, I did"
#
# (The values 1 and 4 — not 1 and 2 — are deliberate; this is how
# the experiment.js encodes them, likely to avoid scale midpoint
# confusion.)
#
# Preregistered criterion: Exclude participants who respond "No, I
# didn't" (seriousness = 1).
# ──────────────────────────────────────────────────────────────────────

cat("Step 4: Self-reported non-seriousness\n")
apply_exclusion(
  condition = df$seriousness == "1",
  label = "4. Self-reported non-serious"
)
```

```{r step-5}
# ──────────────────────────────────────────────────────────────────────
# STEP 5: Self-reported AI use
#
# Participants were asked: "Did you use any AI tool (such as ChatGPT,
# Claude, or similar) to help answer any of the questions in this
# study? Your answer will NOT affect your payment in any way."
# Response options:
#   1 = "Yes"
#   2 = "No"
#
# Preregistered criterion: Exclude participants who respond "Yes"
# (ai_usage = 1).
# ──────────────────────────────────────────────────────────────────────

cat("Step 5: Self-reported AI use\n")
apply_exclusion(
  condition = df$ai_usage == "1",
  label = "5. Self-reported AI use"
)
```

```{r step-6}
# ──────────────────────────────────────────────────────────────────────
# STEP 6: Non-consent or device failure
#
# The experiment logic prevents participants from continuing if they:
#   - Do not consent (consent != 1)
#   - Report using a phone or tablet (device_self_report = 2)
#
# Because the experiment stops them before they can complete, these
# participants should not appear in the Pavlovia data at all. We
# check anyway as a safeguard.
# ──────────────────────────────────────────────────────────────────────

cat("Step 6: Non-consent or device failure\n")
apply_exclusion(
  condition = df$consent != "1" | df$device_self_report == "2",
  label = "6. Non-consent or wrong device"
)
```

```{r step-7}
# ──────────────────────────────────────────────────────────────────────
# STEP 7: Total completion time
#
# Preregistered criterion: Exclude participants who complete the
# entire experiment in under 3 minutes (180,000 ms). The experiment
# includes multiple pages with enforced minimum display times (7s
# each for most question pages), two SPR trials, and several open-
# ended response fields, so a genuine completion in under 3 minutes
# is implausible.
# ──────────────────────────────────────────────────────────────────────

cat("Step 7: Completion time\n")

df$total_duration_ms <- as.numeric(df$total_duration_ms)

apply_exclusion(
  condition = !is.na(df$total_duration_ms) & df$total_duration_ms < 180000,
  label = "7. Completion time < 3 minutes"
)

# Descriptive: show completion time distribution for remaining
remaining <- df[!df$excluded, ]
dur_min <- remaining$total_duration_ms / 60000
cat(sprintf("\n  Completion time — median: %.1f min, range: [%.1f, %.1f]\n",
            median(dur_min, na.rm = TRUE),
            min(dur_min, na.rm = TRUE),
            max(dur_min, na.rm = TRUE)))
```


# 4. Exclusion Summary

```{r exclusion-summary}
# ──────────────────────────────────────────────────────────────────────
# EXCLUSION CASCADE
# Summarize how many were excluded at each step.
# ──────────────────────────────────────────────────────────────────────

cat("═══════════════════════════════════════════════════\n")
cat("  EXCLUSION CASCADE\n")
cat("═══════════════════════════════════════════════════\n\n")

cat(sprintf("  Starting N:                        %d\n", n_start))

# Count by step
step_counts <- table(df$exclusion_step[df$excluded])
for (step in sort(names(step_counts))) {
  cat(sprintf("  %-42s -%d\n", step, step_counts[step]))
}

cat(sprintf("  %s\n", paste(rep("─", 50), collapse = "")))
cat(sprintf("  Total excluded:                    %d\n", sum(df$excluded)))
cat(sprintf("  Final analytic N:                  %d\n", sum(!df$excluded)))
cat(sprintf("  Exclusion rate:                    %.1f%%\n", 
            mean(df$excluded) * 100))
cat("═══════════════════════════════════════════════════\n")
```


# 5. Analytic Sample Descriptives

```{r analytic-sample}
# ──────────────────────────────────────────────────────────────────────
# Quick descriptives of the analytic sample, for comparison with the
# preregistered expectations and the pilot.
# ──────────────────────────────────────────────────────────────────────

analytic <- df[!df$excluded, ]

cat("ANALYTIC SAMPLE DESCRIPTIVES\n\n")

# Interpretation distribution
cat("Interpretation distribution:\n")
interp_table <- table(analytic$interpretation)
for (val in sort(names(interp_table))) {
  label <- c("1" = "Conditional", "2" = "Categorical", "3" = "Neither")[val]
  cat(sprintf("  %s: %d (%.1f%%)\n", label, interp_table[val],
              interp_table[val] / nrow(analytic) * 100))
}

# Sentence distribution
cat("\nSentence distribution:\n")
sent_table <- table(analytic$sentence_id)
cat(sprintf("  Sentences represented: %d of 20\n", length(sent_table)))
cat(sprintf("  Min per sentence: %d\n", min(sent_table)))
cat(sprintf("  Max per sentence: %d\n", max(sent_table)))
cat(sprintf("  Mean per sentence: %.1f\n", mean(sent_table)))

# Flag any sentences with very few observations
low_n <- sent_table[sent_table < 3]
if (length(low_n) > 0) {
  cat("\n  WARNING: Sentences with fewer than 3 observations:\n")
  for (i in seq_along(low_n)) {
    cat(sprintf("    Sentence %s: %d observations\n", names(low_n)[i], low_n[i]))
  }
}

# Rewrite-first counterbalancing
cat("\nRewrite-first counterbalancing:\n")
rf_table <- table(analytic$rewrite_first)
for (val in names(rf_table)) {
  cat(sprintf("  rewrite_first = %s: %d (%.1f%%)\n", val, rf_table[val],
              rf_table[val] / nrow(analytic) * 100))
}

# Demographics overview
cat("\nDemographics:\n")
cat(sprintf("  Age — median: %s, range: [%s, %s]\n",
            median(as.numeric(analytic$age), na.rm = TRUE),
            min(as.numeric(analytic$age), na.rm = TRUE),
            max(as.numeric(analytic$age), na.rm = TRUE)))

gender_table <- table(analytic$gender)
gender_labels <- c("1" = "Male", "2" = "Female", "3" = "Nonbinary", 
                    "4" = "Other", "5" = "Prefer not to say")
for (val in sort(names(gender_table))) {
  cat(sprintf("  Gender %s (%s): %d\n", val, 
              ifelse(val %in% names(gender_labels), gender_labels[val], "Unknown"),
              gender_table[val]))
}
```


# 6. Save Analytic Dataset

```{r save}
# ──────────────────────────────────────────────────────────────────────
# SAVE
# We save the full dataset (including excluded participants) with the
# exclusion flags, AND a filtered version with only the analytic 
# sample. Both are useful:
#   - Full dataset (with flags): for verifying exclusion decisions
#   - Analytic sample: for Stages 3 and 4
#
# We also drop the cleaning columns (rewrite_clean, sentence_clean)
# since they were only needed for edit distance computation.
# ──────────────────────────────────────────────────────────────────────

# Full dataset with exclusion flags
full_output <- file.path(data_dir, "SA_E1A_merged_with_exclusions.csv")
write.csv(df[, !names(df) %in% c("rewrite_clean", "sentence_clean")], 
          full_output, row.names = FALSE, fileEncoding = "UTF-8")
cat("Saved full dataset with exclusion flags to:\n ", full_output, "\n\n")

# Analytic sample only (excluded participants removed)
analytic_out <- analytic[, !names(analytic) %in% c("rewrite_clean", "sentence_clean",
                                                     "excluded", "exclusion_step")]
write.csv(analytic_out, output_file, row.names = FALSE, fileEncoding = "UTF-8")
cat("Saved analytic sample to:\n ", output_file, "\n")
cat("Dimensions:", nrow(analytic_out), "rows x", ncol(analytic_out), "columns.\n")
```


# Summary

```{r final-summary}
cat("═══════════════════════════════════════════════════\n")
cat("  SA_E1A Stage 2: Exclusions — Complete\n")
cat("═══════════════════════════════════════════════════\n\n")
cat("  Starting N:          ", n_start, "\n")
cat("  Excluded:            ", sum(df$excluded), "\n")
cat("  Analytic N:          ", sum(!df$excluded), "\n")
cat("  Exclusion rate:      ", sprintf("%.1f%%", mean(df$excluded) * 100), "\n\n")
cat("  Outputs:\n")
cat("    Full (with flags): ", full_output, "\n")
cat("    Analytic sample:   ", output_file, "\n\n")
cat("  Next steps:\n")
cat("    Stage 3: Rewrite classification (Python/Colab)\n")
cat("    Stage 4: Hypothesis testing (R)\n")
cat("═══════════════════════════════════════════════════\n")
```
