---
title: "SA_E1A — Stage 1: Data Assembly"
author: "David G. Kamper"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
```

# Overview

This document merges the individual per-participant CSV files produced by Pavlovia
into a single rectangular dataset suitable for analysis. Each participant completed the
SA_E1A experiment and produced one CSV file. Because the experiment assigned each
participant exactly one sentence (between-subjects), the sentence-specific columns are
named with a prefix like `S3_rewrite`, `S10_interpretation`, etc. This script detects
each participant's sentence ID and renames those columns to generic names (e.g.,
`rewrite`, `interpretation`) so all participants share a common schema.

The output is a single CSV file: **`SA_E1A_merged.csv`**.

No exclusions are applied here — that happens in Stage 2. This script only assembles
and verifies the raw data.


# 1. Setup and Configuration

```{r config}
# ──────────────────────────────────────────────────────────────────────
# DATA DIRECTORY
# This is the path to the folder containing all individual participant
# CSV files downloaded from Pavlovia. Each file is named like:
#   data_SA_E1A_{prolific_pid}_SESSION_{datetime}.csv
# ──────────────────────────────────────────────────────────────────────

data_dir <- "/Users/dgkamper/Library/CloudStorage/GoogleDrive-dgkamper@gmail.com/My Drive/DGK Lab/Collaborations/BlankLang Lab/DGK Lab - BlankLangLab - Second Amendment Language Dependencies/Data/Study 1/data"

# ──────────────────────────────────────────────────────────────────────
# OUTPUT PATH
# Where the merged dataset will be saved. By default, it goes in the
# same parent directory as the data folder.
# ──────────────────────────────────────────────────────────────────────

output_dir <- dirname(data_dir)
output_file <- file.path(output_dir, "SA_E1A_merged.csv")
```

```{r libraries}
# ──────────────────────────────────────────────────────────────────────
# LIBRARIES
# Only base R and stringr are needed for this stage. We avoid loading
# heavy packages that aren't required yet.
# ──────────────────────────────────────────────────────────────────────

library(stringr)  # for str_replace, str_detect
```


# 2. Sentence Metadata

The experiment uses 20 sentences, each with Second Amendment–style syntax. The
sentence metadata below is transcribed directly from `experiment.js` (the `S` array,
lines 199–219). We need this because the classifier in Stage 3 requires the original
sentence text alongside each participant's rewrite.

The `full()` function in experiment.js constructs the sentence as:
`subject + ", " + being + ", " + right + ", " + must + "."`

```{r sentence-metadata}
# ──────────────────────────────────────────────────────────────────────
# SENTENCE DATA
# Each row corresponds to one of the 20 experimental sentences.
# "full_sentence" is the exact string participants saw on screen, 
# constructed the same way experiment.js builds it with the full() 
# helper function: subject + ", " + being + ", " + right + ", " + must + "."
#
# "no_prefix_sentence" is the version with the prefatory clause removed,
# constructed the same way experiment.js builds it with noPrefix():
# Capitalize first letter of right + ", " + must + "."
# Participants saw this version on the Removal Impact question.
# ──────────────────────────────────────────────────────────────────────

sentences <- data.frame(
  sentence_id = 1:20,
  
  subject = c(
    "A well-kept archive",
    "A respected cultural tradition",
    "A properly organized marketplace",
    "A well-supplied food reserve",
    "A consistently monitored weather station",
    "A structurally sound dam",
    "A tended field",
    "A replenished pharmacy",
    "A functional kitchen",
    "A community garden plot",
    "A dry basement",
    "A marked parking area",
    "A serviceable harbor",
    "A functioning legal system",
    "A free press institution",
    "A representative town council",
    "A reliable postal network",
    "A working public bus system",
    "A consumer co-op",
    "A trained group of medical responders"
  ),
  
  being = c(
    "being necessary for historical preservation",
    "being necessary for fostering unity within a community",
    "being necessary for economic exchange",
    "being necessary for food security",
    "being necessary for the prediction of dangerous storms",
    "being necessary for protection from rising waters",
    "being necessary for seasonal harvest",
    "being necessary for the treatment of illness",
    "being necessary for meal preparation",
    "being necessary for growing vegetables",
    "being necessary for storing household goods",
    "being necessary for the storage of vehicles",
    "being necessary for access to water",
    "being necessary for the administration of justice",
    "being necessary for an informed public",
    "being necessary for local governance",
    "being necessary for civic communication",
    "being necessary for urban mobility",
    "being necessary for fair access to goods",
    "being necessary for public health emergencies"
  ),
  
  right = c(
    "the right of people to access and manage records",
    "the right of people to pass down and share knowledge",
    "the right of people to sell and trade goods",
    "the right of people to store and preserve essential provisions",
    "the right of people to prepare for and respond to natural threats",
    "the right of people to defend their land from flood",
    "the right of people to cultivate and work the land",
    "the right of people to prepare and dispense medicine",
    "the right of people to assemble and equip personal spaces for cooking",
    "the right of people to plant and cultivate their own produce",
    "the right of people to dig and waterproof spaces below the ground",
    "the right of people to park near their homes",
    "the right of people to set anchors along the shore",
    "the right of people to assemble and deliberate as jurors",
    "the right of people to publish and distribute information",
    "the right of people to assemble and deliberate on civic matters",
    "the right of people to send and receive correspondence",
    "the right of people to plan and operate shared transportation",
    "the right of people to pool purchasing power and govern enterprises jointly",
    "the right of people to study and practice emergency care"
  ),
  
  must = c(
    "must not be infringed",
    "must not be infringed",
    "must not be revoked",
    "must not be denied",
    "must not be violated",
    "must not be revoked",
    "must not be infringed",
    "must not be violated",
    "must not be infringed",
    "must not be infringed",
    "must not be revoked",
    "must not be infringed",
    "must not be revoked",
    "must not be hindered",
    "must not be impeded",
    "must not be violated",
    "must not be disrupted",
    "must not be revoked",
    "must not be violated",
    "shall not be infringed"   # Note: sentence 20 uses "shall" not "must"
  ),
  
  stringsAsFactors = FALSE
)

# ──────────────────────────────────────────────────────────────────────
# Construct the full sentence exactly as experiment.js does:
#   full(s) = s.subject + ", " + s.being + ", " + s.right + ", " + s.must + "."
# ──────────────────────────────────────────────────────────────────────

sentences$full_sentence <- paste0(
  sentences$subject, ", ",
  sentences$being, ", ",
  sentences$right, ", ",
  sentences$must, "."
)

# ──────────────────────────────────────────────────────────────────────
# Construct the no-prefix version as experiment.js does:
#   noPrefix(s) = capitalize(s.right) + ", " + s.must + "."
# This is shown on the Removal Impact question.
# ──────────────────────────────────────────────────────────────────────

sentences$no_prefix_sentence <- paste0(
  toupper(substr(sentences$right, 1, 1)),
  substr(sentences$right, 2, nchar(sentences$right)),
  ", ", sentences$must, "."
)

# Quick sanity check: print sentence 10 as a verification
cat("Sentence 10 (verification):\n")
cat(sentences$full_sentence[10], "\n")
```


# 3. Read and Merge Individual CSVs

```{r read-files}
# ──────────────────────────────────────────────────────────────────────
# LIST ALL CSV FILES
# Each participant produces exactly one CSV file. We list them all and
# will process each one individually.
# ──────────────────────────────────────────────────────────────────────

csv_files <- list.files(data_dir, pattern = "\\.csv$", full.names = TRUE)
cat("Found", length(csv_files), "CSV files.\n")
```

```{r define-merge-function}
# ──────────────────────────────────────────────────────────────────────
# FUNCTION: read_and_normalize_one_file
#
# This function reads a single participant CSV and normalizes the
# sentence-specific column names to generic names.
#
# The challenge: Pavlovia records sentence-specific columns with 
# prefixes like S3_, S10_, S17_, etc. Each file only has columns for
# the sentence that participant was assigned. We need to:
#   1. Detect which sentence ID this participant received
#   2. Strip the "S{id}_" prefix from those columns
#   3. Return a data frame with consistent column names
#
# CONDITIONAL COLUMNS:
# - "istrue": Only recorded when interpretation = 1 (conditional) or
#   3 (neither). For categorical interpreters (interpretation = 2),
#   the experiment skips this question entirely, so the column does
#   not exist in their CSV. We set it to NA for those participants.
#
# - "attn1_retry" and "attn1_retry_pass": Only recorded when a
#   participant failed the first attention check. Most participants
#   pass on the first try and don't have these columns. We set them
#   to NA when absent.
# ──────────────────────────────────────────────────────────────────────

read_and_normalize_one_file <- function(filepath) {
  
  # Read the CSV. Each file has exactly one data row (one participant).
  df <- read.csv(filepath, stringsAsFactors = FALSE, check.names = FALSE)
  
  # Basic validation: should have exactly 1 row
  if (nrow(df) != 1) {
    warning(paste("File has", nrow(df), "rows (expected 1):", basename(filepath)))
    return(NULL)
  }
  
  # Get this participant's sentence ID from the "sentence_id" column
  sid <- as.character(df$sentence_id)
  
  # Validate that the sentence ID is between 1 and 20
  if (!sid %in% as.character(1:20)) {
    warning(paste("Invalid sentence_id:", sid, "in", basename(filepath)))
    return(NULL)
  }
  
  # ── Identify sentence-specific columns ──
  # These are columns that match the pattern "S{id}_something"
  # For example, if sentence_id = 10: S10_rewrite, S10_interpretation, etc.
  prefix <- paste0("S", sid, "_")
  sid_cols <- names(df)[str_detect(names(df), paste0("^", prefix))]
  
  # ── Rename sentence-specific columns to generic names ──
  # "S10_rewrite" becomes "rewrite", "S10_interpretation" becomes "interpretation", etc.
  new_names <- str_replace(sid_cols, paste0("^", prefix), "")
  names(df)[match(sid_cols, names(df))] <- new_names
  
  # ── Handle the conditional "istrue" column ──
  # If interpretation was 2 (categorical), this column doesn't exist.
  # We add it as NA so all participants have the same columns.
  if (!"istrue" %in% names(df)) {
    df$istrue <- NA
  }
  
  # ── Handle the conditional attention check retry columns ──
  # Most participants pass attn1 on the first try and don't have these.
  if (!"attn1_retry" %in% names(df)) {
    df$attn1_retry <- NA
  }
  if (!"attn1_retry_pass" %in% names(df)) {
    df$attn1_retry_pass <- NA
  }
  
  # ── Add source filename for traceability ──
  df$source_file <- basename(filepath)
  
  return(df)
}
```

```{r merge-all-files}
# ──────────────────────────────────────────────────────────────────────
# PROCESS ALL FILES
# Read each file, normalize column names, and combine into one data frame.
# We track any files that fail to process and report them.
# ──────────────────────────────────────────────────────────────────────

all_dfs <- list()
failed_files <- character(0)

for (i in seq_along(csv_files)) {
  result <- tryCatch(
    read_and_normalize_one_file(csv_files[i]),
    error = function(e) {
      warning(paste("Error reading file", basename(csv_files[i]), ":", e$message))
      return(NULL)
    }
  )
  
  if (!is.null(result)) {
    all_dfs[[length(all_dfs) + 1]] <- result
  } else {
    failed_files <- c(failed_files, basename(csv_files[i]))
  }
}

cat("Successfully processed:", length(all_dfs), "files.\n")

if (length(failed_files) > 0) {
  cat("FAILED files:\n")
  for (f in failed_files) cat("  -", f, "\n")
} else {
  cat("No failures.\n")
}

# ──────────────────────────────────────────────────────────────────────
# COMBINE INTO ONE DATA FRAME
# All data frames should now have identical column names. We use
# rbind to stack them. If column mismatches remain, this will error
# informatively.
# ──────────────────────────────────────────────────────────────────────

# First, verify all data frames have the same column set
col_sets <- lapply(all_dfs, names)
reference_cols <- sort(col_sets[[1]])

mismatched <- which(sapply(col_sets, function(x) !identical(sort(x), reference_cols)))

if (length(mismatched) > 0) {
  cat("\nWARNING: Column mismatch in", length(mismatched), "file(s).\n")
  for (idx in mismatched) {
    extra <- setdiff(names(all_dfs[[idx]]), names(all_dfs[[1]]))
    missing <- setdiff(names(all_dfs[[1]]), names(all_dfs[[idx]]))
    cat("  File", idx, "- extra:", paste(extra, collapse = ", "), 
        "| missing:", paste(missing, collapse = ", "), "\n")
  }
  
  # ── Harmonize columns across all data frames ──
  # Use the union of all column names so no data is lost. Missing
  # columns in any given data frame get filled with NA.
  all_col_names <- unique(unlist(col_sets))
  
  all_dfs <- lapply(all_dfs, function(df) {
    missing_cols <- setdiff(all_col_names, names(df))
    for (col in missing_cols) {
      df[[col]] <- NA
    }
    # Reorder columns to match
    df[, all_col_names]
  })
}

# Stack all rows
merged <- do.call(rbind, all_dfs)
rownames(merged) <- NULL

cat("\nMerged dataset:", nrow(merged), "participants x", ncol(merged), "columns.\n")
```


# 4. Join Sentence Metadata

```{r join-metadata}
# ──────────────────────────────────────────────────────────────────────
# JOIN SENTENCE TEXT
# We attach the original sentence text, its components, and the
# no-prefix version to each participant's row. This is needed for:
#   - Stage 3 (classifier needs the original sentence + rewrite)
#   - Any analysis that references the sentence content
#
# We join on sentence_id after ensuring both are numeric.
# ──────────────────────────────────────────────────────────────────────

merged$sentence_id <- as.integer(merged$sentence_id)

merged <- merge(merged, sentences, by = "sentence_id", all.x = TRUE)

# Verify: no participants should have missing sentence metadata
missing_sentence <- sum(is.na(merged$full_sentence))
if (missing_sentence > 0) {
  warning(paste(missing_sentence, "participants have no matching sentence metadata!"))
} else {
  cat("All participants successfully matched to sentence metadata.\n")
}
```


# 5. Flag Test Runs

```{r flag-test-runs}
# ──────────────────────────────────────────────────────────────────────
# FLAG TEST / PILOT RUNS
# During platform testing, the experimenter may have completed the
# experiment using a placeholder PID (e.g., "TEST"). These files get
# saved to Pavlovia alongside real participant data. We flag them here
# so they can be identified, but we do NOT remove them — exclusions
# happen in Stage 2.
#
# Known indicator: prolific_pid = "TEST" and missing captcha/SPR
# columns (the screening tasks were added after the initial test run).
# ──────────────────────────────────────────────────────────────────────

test_runs <- merged$prolific_pid %in% c("TEST", "test", "")
cat("Flagged as test runs:", sum(test_runs), "\n")
if (sum(test_runs) > 0) {
  cat("  PIDs:", paste(merged$prolific_pid[test_runs], collapse = ", "), "\n")
  cat("  Source files:", paste(merged$source_file[test_runs], collapse = ", "), "\n")
  cat("  (These will be excluded in Stage 2.)\n")
}
```


# 6. Verification Checks

Before saving, we run a series of verification checks.

```{r verify-dimensions}
# ──────────────────────────────────────────────────────────────────────
# CHECK 1: Row count matches file count
# Each file should produce exactly one row.
# ──────────────────────────────────────────────────────────────────────

cat("Files read:", length(all_dfs), "\n")
cat("Rows in merged:", nrow(merged), "\n")

if (nrow(merged) != length(all_dfs)) {
  warning("Row count does not match file count!")
}
```

```{r verify-duplicates}
# ──────────────────────────────────────────────────────────────────────
# CHECK 2: No duplicate Prolific PIDs
# Each participant should appear exactly once. Duplicates would mean
# either the same person completed the experiment twice, or we have
# duplicate CSV files.
# ──────────────────────────────────────────────────────────────────────

pid_counts <- table(merged$prolific_pid)
duplicates <- pid_counts[pid_counts > 1]

if (length(duplicates) > 0) {
  cat("WARNING: Duplicate Prolific PIDs found:\n")
  print(duplicates)
} else {
  cat("No duplicate Prolific PIDs. Each participant appears once.\n")
}
```

```{r verify-sentence-distribution}
# ──────────────────────────────────────────────────────────────────────
# CHECK 3: Sentence distribution
# The experiment assigns sentences via Math.floor(Math.random()*20),
# so assignment is random (not counterbalanced). We expect uneven
# counts but all 20 sentences should be represented.
# ──────────────────────────────────────────────────────────────────────

sentence_dist <- table(merged$sentence_id)
cat("Sentence distribution:\n")
print(sentence_dist)
cat("\nSentences represented:", length(sentence_dist), "of 20\n")
cat("Min per sentence:", min(sentence_dist), "\n")
cat("Max per sentence:", max(sentence_dist), "\n")
cat("Mean per sentence:", round(mean(sentence_dist), 1), "\n")
```

```{r verify-key-columns}
# ──────────────────────────────────────────────────────────────────────
# CHECK 4: Key columns are non-missing
# The columns that are critical for analysis should have values for
# every participant (except istrue, which is legitimately NA for
# categorical interpreters).
# ──────────────────────────────────────────────────────────────────────

key_cols <- c("prolific_pid", "sentence_id", "interpretation", "confidence",
              "rewrite", "addcommas", "remove", "and",
              "comma_3v2", "comma_3v1", "comma_2v1", "distinguish",
              "captcha_pass", "attn1_pass", "seriousness", "ai_usage",
              "total_duration_ms", "rewrite_first")

cat("Missing values in key columns:\n")
for (col in key_cols) {
  n_missing <- sum(is.na(merged[[col]]) | merged[[col]] == "")
  if (n_missing > 0) {
    cat("  ", col, ":", n_missing, "missing\n")
  }
}
cat("(No output above = no missing values in key columns.)\n")
```

```{r verify-interpretation}
# ──────────────────────────────────────────────────────────────────────
# CHECK 5: Interpretation distribution
# Quick look at the primary DV to make sure values are as expected.
# Values should be 1 (conditional), 2 (categorical), or 3 (neither).
# ──────────────────────────────────────────────────────────────────────

cat("Interpretation distribution:\n")
interp_table <- table(merged$interpretation, useNA = "ifany")
print(interp_table)

cat("\nAs percentages:\n")
print(round(prop.table(interp_table) * 100, 1))
```

```{r verify-istrue-logic}
# ──────────────────────────────────────────────────────────────────────
# CHECK 6: Verify istrue is present/absent as expected
# The istrue question only appears when interpretation = 1 or 3.
# So istrue should be NA when interpretation = 2, and non-NA otherwise.
# ──────────────────────────────────────────────────────────────────────

# Participants who chose categorical (2) should have NA for istrue
cat_no_istrue <- sum(merged$interpretation == "2" & is.na(merged$istrue))
cat_has_istrue <- sum(merged$interpretation == "2" & !is.na(merged$istrue))

# Participants who chose conditional (1) or neither (3) should have istrue
other_has_istrue <- sum(merged$interpretation %in% c("1", "3") & !is.na(merged$istrue))
other_no_istrue <- sum(merged$interpretation %in% c("1", "3") & is.na(merged$istrue))

cat("Categorical (interp=2): istrue is NA =", cat_no_istrue, ", has value =", cat_has_istrue, "\n")
cat("Conditional/Neither (interp=1,3): istrue has value =", other_has_istrue, ", is NA =", other_no_istrue, "\n")

if (cat_has_istrue > 0 | other_no_istrue > 0) {
  warning("istrue column logic does not match expectations!")
} else {
  cat("istrue column logic verified: matches experiment flow.\n")
}
```


# 7. Column Inventory

```{r column-inventory}
# ──────────────────────────────────────────────────────────────────────
# Print all column names in the final merged dataset, grouped by
# category, for reference.
# ──────────────────────────────────────────────────────────────────────

cat("FINAL COLUMN INVENTORY (", ncol(merged), " columns):\n\n")

id_cols <- c("prolific_pid", "prolific_study_id", "prolific_session_id",
             "sentence_id", "source_file")
cat("Identifiers:", paste(id_cols[id_cols %in% names(merged)], collapse = ", "), "\n\n")

experiment_cols <- c("rewrite_first", "start_time", "end_time", "total_duration_ms",
                     "device_type", "user_agent", "fullscreen_entered", 
                     "device_self_report", "screen_width", "screen_height",
                     "page_timestamps")
cat("Experiment metadata:", paste(experiment_cols[experiment_cols %in% names(merged)], collapse = ", "), "\n\n")

screening_cols <- c("captcha_response", "captcha_pass", 
                    "spr1_rts", "spr1_mean_rt", "spr1_min_rt", "spr1_understood",
                    "spr2_rts", "spr2_mean_rt", "spr2_min_rt", "spr2_understood",
                    "consent", "date_of_consent",
                    "attn1", "attn1_pass", "attn1_retry", "attn1_retry_pass")
cat("Screening / AI detection:", paste(screening_cols[screening_cols %in% names(merged)], collapse = ", "), "\n\n")

dv_cols <- c("rewrite", "addcommas", "interpretation", "confidence", "istrue",
             "remove", "and", "comma_3v2", "comma_3v1", "comma_2v1", "distinguish")
cat("Dependent variables:", paste(dv_cols[dv_cols %in% names(merged)], collapse = ", "), "\n\n")

attn2_cols <- c("police1", "police2", "police3")
cat("Attention check 2:", paste(attn2_cols[attn2_cols %in% names(merged)], collapse = ", "), "\n\n")

demo_cols <- c("gender", "age", "race", "country", "native_language",
               "english_exposure_age", "other_languages", "english_frequency",
               "potential_conflicts", "potential_conflicts_explain", "gun_control")
cat("Demographics:", paste(demo_cols[demo_cols %in% names(merged)], collapse = ", "), "\n\n")

debrief_cols <- c("ai_usage", "seriousness", "quiet_environment", "distracted",
                  "feedback_procedure", "feedback_other")
cat("Debrief:", paste(debrief_cols[debrief_cols %in% names(merged)], collapse = ", "), "\n\n")

sentence_meta_cols <- c("subject", "being", "right", "must", 
                        "full_sentence", "no_prefix_sentence")
cat("Sentence metadata (joined):", paste(sentence_meta_cols[sentence_meta_cols %in% names(merged)], collapse = ", "), "\n\n")
```


# 8. Save

```{r save}
# ──────────────────────────────────────────────────────────────────────
# SAVE THE MERGED DATASET
# Write as CSV with UTF-8 encoding. Row names are excluded.
# This file is the input for all subsequent stages:
#   - Stage 2 (exclusions)
#   - Stage 3 (rewrite classification, in Python/Colab)
#   - Stage 4 (hypothesis testing)
# ──────────────────────────────────────────────────────────────────────

write.csv(merged, output_file, row.names = FALSE, fileEncoding = "UTF-8")
cat("Saved merged dataset to:\n", output_file, "\n")
cat("Dimensions:", nrow(merged), "rows x", ncol(merged), "columns.\n")
```


# Summary

```{r summary}
# ──────────────────────────────────────────────────────────────────────
# FINAL SUMMARY
# ──────────────────────────────────────────────────────────────────────

cat("═══════════════════════════════════════════════════\n")
cat("  SA_E1A Stage 1: Data Assembly — Complete\n")
cat("═══════════════════════════════════════════════════\n\n")
cat("  Total participants:        ", nrow(merged), "\n")
cat("  Sentences represented:     ", length(unique(merged$sentence_id)), "of 20\n")
cat("  Failed files:              ", length(failed_files), "\n")
cat("  Duplicate PIDs:            ", sum(pid_counts > 1), "\n")
cat("  Interpretation breakdown:  \n")
cat("    Categorical (2):         ", sum(merged$interpretation == "2"), 
    " (", round(mean(merged$interpretation == "2") * 100, 1), "%)\n", sep = "")
cat("    Conditional (1):         ", sum(merged$interpretation == "1"), 
    " (", round(mean(merged$interpretation == "1") * 100, 1), "%)\n", sep = "")
cat("    Neither (3):             ", sum(merged$interpretation == "3"), 
    " (", round(mean(merged$interpretation == "3") * 100, 1), "%)\n", sep = "")
cat("\n  Output: ", output_file, "\n")
cat("  Next step: Stage 2 (Exclusions)\n")
cat("═══════════════════════════════════════════════════\n")
```
